---
title: "04_differential_abundance_ANCOMBC"
author: "Zulema Carracedo Lorenzo"
date: "2025-03-12"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Load packages
```{r}
library(phyloseq)
library(ggplot2)
library(dplyr)
library(tidyr)
library(vegan)
library(UpSetR)
library(purrr)
library(tibble)
library(tidyr)
library(usefun)
library(writexl)

#Differential abundance
library(metagMisc) #for CSS transformation & prevalence filter
library(metagenomeSeq)
library(DESeq2) #differential abundance
library(zinbwave) #zero-inflated
library(ANCOMBC) #ANCOM
#library(GUniFrac) #ZicoSeq & rarefaction PROBLEM
library(ALDEx2) #Aldex2
library(Maaslin2) #MaAslin2
#library(corncob) #not able to install

#Heatmaps
library(ComplexHeatmap)

```

#Load phyloseq object
```{r}
#For differential abundance analysis we use an unnormalize phyloseq object

load("./5_Brassica_oleracea_exp/r_objects/ps_filtered.RData") 

#We will only use the samples from the root soil

ps_filtered_root <-  subset_samples(ps_filtered, Sample_type %in% "root_associated")
ps_filtered_root <- prune_taxa(taxa_sums(ps_filtered_root) >0, ps_filtered_root)

rm(ps_filtered)
```

#Define pairwise comparisons
```{r}

#We need to divide our data into WW and DR

ps_WW_root <- subset_samples(ps_filtered_root, water_regime %in% "WW")
ps_WW_root <- prune_taxa(taxa_sums(ps_WW_root) >0, ps_WW_root)

ps_DR_root <- subset_samples(ps_filtered_root, water_regime %in% "DR")
ps_DR_root <- prune_taxa(taxa_sums(ps_DR_root) >0, ps_DR_root)


```



# Filtering low prevalence taxa

```{r}
ps_filtered_root #48,027 ASV

### WW ###
ps_WW_root #34,464 ASVs

#First, check how many zero's
zero_percentage <- function(x){
   df <- data.frame(otu_table(x))
   zero_percentage <- (sum(colSums(df == 0))) / (nrow(df) * ncol(df))
   return(zero_percentage)
}

zero_percentage(ps_WW_root) #84.31%

# Now, drop taxa present in at least 20% of the samples
ps_WW_root_f <- metagMisc::phyloseq_filter_prevalence(ps_WW_root, prev.trh = 0.20, #at least 20% of samples
                                          abund.trh = NULL) #no filtering by abundance 

#check
ps_WW_root_f #6,846 ASVs

zero_percentage(ps_WW_root_f) # 43.75%


### DR ###
ps_DR_root #29,367 ASVs

#First, check how many zero's
zero_percentage(ps_DR_root) #83.15%

# Now, drop taxa present in at least 20% of the samples
ps_DR_root_f <- metagMisc::phyloseq_filter_prevalence(ps_DR_root, prev.trh = 0.20, #at least 20% of samples
                                          abund.trh = NULL) #no filtering by abundance 

#check
ps_DR_root_f #6,122 ASVs

zero_percentage(ps_DR_root_f) # 41.34%

rm(ps_WW_root,ps_DR_root)


```


# 1. ANCOM-BC

ANCOM-BC (Analysis of Compositions of Microbiomes with Bias Correction) is a method that takes into account the compositionality of microbiome data by using an additive log-ratio transformation. A ratio is calculated for each taxon with a reference taxon with low variance present across all samples. The differences between the ratios are compared and tested statistically to determine whether a taxon is DA between two conditions or not. It is considered a stringent method. Input is non-rarefied data. 

This is a good explanation of ANCOM-BC : https://github.com/FrederickHuangLin/ANCOM-BC-Code-Archive

##WW
```{r}

# Note that by default, levels of a categorical variable in R are sorted alphabetically. In this case, the reference level for `treatment` will be`CHA0`. To manually change the reference level, use:

ps_WW_root_f@sam_data$treatment = factor(ps_WW_root_f@sam_data$treatment, levels = c("Control", "WCS417", "WCS315", "CHA0"))

# You can verify the change by checking:
levels(ps_WW_root_f@sam_data$treatment)

#Calculate ANCOm-BC

  
  set.seed(123)
  ancom <- ancombc(phyloseq = ps_WW_root_f, 
                      formula = "treatment", 
                      p_adj_method = "fdr", #same as "BH" Benjamini & Hochberg
                      zero_cut = 0.9, #we are removing ASVs that have more than 90% of 0s
                      lib_cut = 0, 
                      group = "treatment", 
                      struc_zero = TRUE, #with structural zero's
                      neg_lb = FALSE, #classifies structural zero's if it's zero and not its asymptotic lower bound 
                      tol = 1e-5,               
                      max_iter = 100, 
                      conserve = TRUE, 
                      alpha = 0.05)

  save(ancom, file="./5_Brassica_oleracea_exp/r_objects/ancom_bc_root_WW.RData")
  load(file="./5_Brassica_oleracea_exp/r_objects/ancom_bc_root_WW.RData")
  
  #get results from ANCOM-BC into df
 df_ancom_WW <- data.frame(
  ASV = row.names(ancom$res$beta), #get ASV names
  lfc = unlist(ancom$res$beta),
  se = unlist(ancom$res$se),
  W = unlist(ancom$res$W),
  p_val = unlist(ancom$res$p_val),
  q_val = unlist(ancom$res$q_val),
  diff_abn = unlist(ancom$res$diff_abn))
 
 #Row names to column and fix teatment names
 
 df_ancom_WW <- rownames_to_column(df_ancom_WW)
 
 df_ancom_WW <- df_ancom_WW %>% rename("rowname" = "treatment")
 
df_ancom_WW <- df_ancom_WW %>%
   mutate(Treatment =
            case_when(grepl("WCS417", df_ancom_WW$treatment) ~  "WCS417",
                      grepl("WCS315", df_ancom_WW$treatment) ~  "WCS315",
                     grepl("CHA0", df_ancom_WW$treatment) ~  "CHA0" 
                      ))

df_ancom_WW <- df_ancom_WW[,2:9] #removes wrong treatment column

#store in a list a singel dataframe per treatment

df_ancom_WW_417 <- df_ancom_WW %>% filter(Treatment == "WCS417")
df_ancom_WW_315 <- df_ancom_WW %>% filter(Treatment == "WCS315")
df_ancom_WW_CHA0 <- df_ancom_WW %>% filter(Treatment == "CHA0")

df_ancom_WW_root_l <- list(df_ancom_WW_417, df_ancom_WW_315, df_ancom_WW_CHA0)
names(df_ancom_WW_root_l) <- c("WCS417", "WCS315", "CHA0")

rm(df_ancom_WW_417, df_ancom_WW_315, df_ancom_WW_CHA0) 

#place ASVs as rownames, for some reason didn't allow me to do it with lapply

  df_ancom_WW_root_l$WCS417 <- df_ancom_WW_root_l$WCS417 %>% 
              remove_rownames()%>%
              column_to_rownames(var="ASV")

  df_ancom_WW_root_l$WCS315 <- df_ancom_WW_root_l$WCS315 %>% 
              remove_rownames()%>%
              column_to_rownames(var="ASV")
  
    df_ancom_WW_root_l$CHA0 <- df_ancom_WW_root_l$CHA0 %>% 
              remove_rownames()%>%
              column_to_rownames(var="ASV")
    
  
  #join df's from ANCOMBC and phyloseq
    
  df_ancom_WW_root_l$WCS417  <- cbind(as(df_ancom_WW_root_l$WCS417, "data.frame"),
                  as(tax_table(ps_WW_root_f)[rownames(df_ancom_WW_root_l$WCS417), ], "matrix"))
  
  df_ancom_WW_root_l$WCS315  <- cbind(as(df_ancom_WW_root_l$WCS315, "data.frame"),
                  as(tax_table(ps_WW_root_f)[rownames(df_ancom_WW_root_l$WCS315), ], "matrix"))
  
  df_ancom_WW_root_l$CHA0  <- cbind(as(df_ancom_WW_root_l$CHA0, "data.frame"),
                  as(tax_table(ps_WW_root_f)[rownames(df_ancom_WW_root_l$CHA0), ], "matrix"))

  
  #keep only those with adjusted p value of <0.05 ("q_val") 

  
  df_ancom_WW_ps_sig_WCS417 <- df_ancom_WW_root_l$WCS417 %>% 
   dplyr::filter(q_val < 0.05)%>%
   mutate(test="3_ancombc") #1045
  
    df_ancom_WW_ps_sig_WCS315 <- df_ancom_WW_root_l$WCS315 %>% 
   dplyr::filter(q_val < 0.05)%>%
   mutate(test="3_ancombc") #627
    
      df_ancom_WW_ps_sig_CHA0 <- df_ancom_WW_root_l$CHA0 %>% 
   dplyr::filter(q_val < 0.05)%>%
   mutate(test="3_ancombc") #629
  
df_ancom_WW_ps_sig_l <- list(df_ancom_WW_ps_sig_WCS417, df_ancom_WW_ps_sig_WCS315, df_ancom_WW_ps_sig_CHA0) #list of lists of significant ASVs detectec by ANCOM by each treatment

names(df_ancom_WW_ps_sig_l) <- c("WCS417", "WCS315", "CHA0")



#
#check how many DA asvs were

sum(df_ancom_WW_root_l$WCS417$diff_abn == "TRUE") #1,045
sum(df_ancom_WW_root_l$WCS315$diff_abn == "TRUE") #627
sum(df_ancom_WW_root_l$CHA0$diff_abn == "TRUE") #629

#Save test results and significan ASV list in a single object

ancombc_WW_l <- list(results= df_ancom_WW_root_l, df_sig_asv_l = df_ancom_WW_ps_sig_l)


##Save
#save(ancom_bc_l, file="./results/r_objects/differential_abundance/1_ancombc_l.RData")
#load(file="./results/r_objects/differential_abundance/1_ancombc_l.RData")

#save list of ASVs per strains
ancombc_WW_WCS417 <- ancombc_WW_l$df_sig_asv_l$WCS417
write_xlsx(ancombc_WW_WCS417, "./5_Brassica_oleracea_exp/results/tables/diff_abundance/ancombc_WW_WCS417.xlsx")

ancombc_WW_WCS315 <- ancombc_WW_l$df_sig_asv_l$WCS315
write_xlsx(ancombc_WW_WCS315, "./5_Brassica_oleracea_exp/results/tables/diff_abundance/ancombc_WW_WCS315.xlsx")

ancombc_WW_CHA0 <- ancombc_WW_l$df_sig_asv_l$CHA0
write_xlsx(ancombc_WW_CHA0, "./5_Brassica_oleracea_exp/results/tables/diff_abundance/ancombc_WW_CHAO.xlsx")

         
```

##DR
```{r}

# Note that by default, levels of a categorical variable in R are sorted alphabetically. In this case, the reference level for `treatment` will be`CHA0`. To manually change the reference level, use:

ps_DR_root_f@sam_data$treatment = factor(ps_DR_root_f@sam_data$treatment, levels = c("Control", "WCS417", "WCS315", "CHA0"))

# You can verify the change by checking:
levels(ps_DR_root_f@sam_data$treatment)

#Calculate ANCOm-BC

  
  set.seed(123)
  ancom <- ancombc(phyloseq = ps_DR_root_f, 
                      formula = "treatment", 
                      p_adj_method = "fdr", #same as "BH" Benjamini & Hochberg
                      zero_cut = 0.9, #we are removing ASVs that have more than 90% of 0s
                      lib_cut = 0, 
                      group = "treatment", 
                      struc_zero = TRUE, #with structural zero's
                      neg_lb = FALSE, #classifies structural zero's if it's zero and not its asymptotic lower bound 
                      tol = 1e-5,               
                      max_iter = 100, 
                      conserve = TRUE, 
                      alpha = 0.05)

  save(ancom, file="./5_Brassica_oleracea_exp/r_objects//ancom_bc_root_DR.RData")
  load(file="./5_Brassica_oleracea_exp/r_objects//ancom_bc_root_DR.RData")
  
  #get results from ANCOM-BC into df
 df_ancom_DR <- data.frame(
  ASV = row.names(ancom$res$beta), #get ASV names
  lfc = unlist(ancom$res$beta),
  se = unlist(ancom$res$se),
  W = unlist(ancom$res$W),
  p_val = unlist(ancom$res$p_val),
  q_val = unlist(ancom$res$q_val),
  diff_abn = unlist(ancom$res$diff_abn))
 
 #Row names to column and fix teatment names
 
 df_ancom_DR <- rownames_to_column(df_ancom_DR)
 
 df_ancom_DR <- df_ancom_DR %>% rename("rowname" = "treatment")
 
df_ancom_DR <- df_ancom_DR %>%
   mutate(Treatment =
            case_when(grepl("WCS417", df_ancom_DR$treatment) ~  "WCS417",
                      grepl("WCS315", df_ancom_DR$treatment) ~  "WCS315",
                     grepl("CHA0", df_ancom_DR$treatment) ~  "CHA0" 
                      ))

df_ancom_DR <- df_ancom_DR[,2:9] #removes wrong treatment column

#store in a list a singel dataframe per treatment

df_ancom_DR_417 <- df_ancom_DR %>% filter(Treatment == "WCS417")
df_ancom_DR_315 <- df_ancom_DR %>% filter(Treatment == "WCS315")
df_ancom_DR_CHA0 <- df_ancom_DR %>% filter(Treatment == "CHA0")

df_ancom_DR_root_l <- list(df_ancom_DR_417, df_ancom_DR_315, df_ancom_DR_CHA0)
names(df_ancom_DR_root_l) <- c("WCS417", "WCS315", "CHA0")

rm(df_ancom_DR_417, df_ancom_DR_315, df_ancom_DR_CHA0) 

#place ASVs as rownames, for some reason didn't allow me to do it with lapply

  df_ancom_DR_root_l$WCS417 <- df_ancom_DR_root_l$WCS417 %>% 
              remove_rownames()%>%
              column_to_rownames(var="ASV")

  df_ancom_DR_root_l$WCS315 <- df_ancom_DR_root_l$WCS315 %>% 
              remove_rownames()%>%
              column_to_rownames(var="ASV")
  
    df_ancom_DR_root_l$CHA0 <- df_ancom_DR_root_l$CHA0 %>% 
              remove_rownames()%>%
              column_to_rownames(var="ASV")
    
  
  #join df's from ANCOMBC and phyloseq
    
  df_ancom_DR_root_l$WCS417  <- cbind(as(df_ancom_DR_root_l$WCS417, "data.frame"),
                  as(tax_table(ps_DR_root_f)[rownames(df_ancom_DR_root_l$WCS417), ], "matrix"))
  
  df_ancom_DR_root_l$WCS315  <- cbind(as(df_ancom_DR_root_l$WCS315, "data.frame"),
                  as(tax_table(ps_DR_root_f)[rownames(df_ancom_DR_root_l$WCS315), ], "matrix"))
  
  df_ancom_DR_root_l$CHA0  <- cbind(as(df_ancom_DR_root_l$CHA0, "data.frame"),
                  as(tax_table(ps_DR_root_f)[rownames(df_ancom_DR_root_l$CHA0), ], "matrix"))

  
  #keep only those with adjusted p value of <0.05 ("q_val") 

  
  df_ancom_DR_ps_sig_WCS417 <- df_ancom_DR_root_l$WCS417 %>% 
   dplyr::filter(q_val < 0.05)%>%
   mutate(test="3_ancombc") #695
  
    df_ancom_DR_ps_sig_WCS315 <- df_ancom_DR_root_l$WCS315 %>% 
   dplyr::filter(q_val < 0.05)%>%
   mutate(test="3_ancombc") #630
    
      df_ancom_DR_ps_sig_CHA0 <- df_ancom_DR_root_l$CHA0 %>% 
   dplyr::filter(q_val < 0.05)%>%
   mutate(test="3_ancombc") #386
  
df_ancom_DR_ps_sig_l <- list(df_ancom_DR_ps_sig_WCS417, df_ancom_DR_ps_sig_WCS315, df_ancom_DR_ps_sig_CHA0) #list of lists of significant ASVs detectec by ANCOM by each treatment

names(df_ancom_DR_ps_sig_l) <- c("WCS417", "WCS315", "CHA0")


#check how many DA asvs were

sum(df_ancom_DR_root_l$WCS417$diff_abn == "TRUE") #695
sum(df_ancom_DR_root_l$WCS315$diff_abn == "TRUE") #630
sum(df_ancom_DR_root_l$CHA0$diff_abn == "TRUE") #386

#Save test results and significan ASV list in a single object

ancombc_DR_l <- list(results= df_ancom_DR_root_l, df_sig_asv_l = df_ancom_DR_ps_sig_l)

#check which ones 
#lapply(ancom_bc_l, function(x) rownames(x$df_sig_asv)) 

#Save
#save(ancom_bc_l, file="./results/r_objects/differential_abundance/1_ancombc_l.RData")
#load(file="./results/r_objects/differential_abundance/1_ancombc_l.RData")

#save list of ASVs per strains
ancombc_DR_WCS417 <- ancombc_DR_l$df_sig_asv_l$WCS417
write_xlsx(ancombc_DR_WCS417, "./5_Brassica_oleracea_exp/results/tables/diff_abundance/ancombc_DR_WCS417.xlsx")

ancombc_DR_WCS315 <- ancombc_DR_l$df_sig_asv_l$WCS315
write_xlsx(ancombc_DR_WCS315, "./5_Brassica_oleracea_exp/results/tables/diff_abundance/ancombc_DR_WCS315.xlsx")

ancombc_DR_CHA0 <- ancombc_DR_l$df_sig_asv_l$CHA0
write_xlsx(ancombc_DR_CHA0, "./5_Brassica_oleracea_exp/results/tables/diff_abundance/ancombc_DR_CHA0.xlsx")

         
```


#Plotting number of ASVs enriched/depleated by 3 strains

```{r}

### well-watered ###
  ancombc_WW_l

WCS417_df <- ancombc_WW_l$df_sig_asv_l$WCS417 %>% 
  rownames_to_column(var = "ASV")

WCS315_df <- ancombc_WW_l$df_sig_asv_l$WCS315 %>% 
  rownames_to_column(var = "ASV")

CHA0_df <- ancombc_WW_l$df_sig_asv_l$CHA0 %>% 
  rownames_to_column(var = "ASV")

#join in a single dataframe
df_combined <- rbind(WCS417_df, WCS315_df, CHA0_df)

#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- WCS417_df$ASV #1045 ASVs
DA_WCS315_vector <- WCS315_df$ASV #627 ASVs
DA_CHA0_vector <- CHA0_df$ASV #629 ASVs

#I want to keep only the ASVs  are differentially abundance under the three treatments when compared to control, so keep only ASVs that are present in the 3 vectors

DA_all_treat_vector <- Reduce(intersect, list(DA_WCS417_vector,DA_WCS315_vector,DA_CHA0_vector)) #261 ASVs

df_combined_WW_f <- df_combined %>%
  filter(ASV %in% DA_all_treat_vector)

#Now let's try to make a plot with both WW and DR data

df_combined_DR_f
df_combined_WW_f <- df_combined_f

### drought ###

    ancombc_DR_l

WCS417_df <- ancombc_DR_l$df_sig_asv_l$WCS417 %>% 
  rownames_to_column(var = "ASV")

WCS315_df <- ancombc_DR_l$df_sig_asv_l$WCS315 %>% 
  rownames_to_column(var = "ASV")

CHA0_df <- ancombc_DR_l$df_sig_asv_l$CHA0 %>% 
  rownames_to_column(var = "ASV")

#join in a single dataframe
df_combined_DR <- rbind(WCS417_df, WCS315_df, CHA0_df)

# What if I select only the ASVs that are significant for the 3 treatments

df_combined_DR

#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- WCS417_df$ASV #695 ASVs
DA_WCS315_vector <- WCS315_df$ASV #630 ASVs
DA_CHA0_vector <- CHA0_df$ASV #386 ASVs

#I want to keep only the ASVs  are differentially abundance under the three treatments when compared to control, so keep only ASVs that are present in the 3 vectors

DA_all_treat_vector <- Reduce(intersect, list(DA_WCS417_vector,DA_WCS315_vector,DA_CHA0_vector)) #386 ASVs

df_combined_DR_f <- df_combined_DR %>%
  filter(ASV %in% DA_all_treat_vector)

### Join well-watered and drought df ###

#add a column for the water regime

df_combined_DR_f <- df_combined_DR_f %>% mutate(water_regime = "DR")
df_combined_WW_f <- df_combined_WW_f %>% mutate(water_regime = "WW")

#join the df

df_combined_f <- rbind(df_combined_DR_f, df_combined_WW_f)

# Optional — remove prefixes like "p__", "f__", etc., for cleaner plot labels
clean_taxonomy <- function(x) {
  x <- trimws(x)  # Remove leading/trailing spaces
  gsub("^\\w__", "", x)  # Remove the prefix
}

# Apply the function with mutate
df_combined_f <- df_combined_f %>%
  mutate(across(c(Phylum, Family, Genus), clean_taxonomy))


#check how many ASVs are depleated/enriched per phylun for each treatment and water regime

# Step 1: Add 'status' column to classify LFC direction
df_summary <- df_combined_f %>%
  mutate(status = ifelse(lfc > 0, "Enriched", "Depleted")) %>%
  group_by(Phylum, Treatment, water_regime, status) %>%
  summarise(n_ASVs = n_distinct(ASV), .groups = "drop") %>%
  # Assign negative sign to depleted counts
  mutate(n_ASVs = ifelse(status == "Depleted", -n_ASVs, n_ASVs))

#plot

df_summary$Treatment <- factor(df_summary$Treatment, c("WCS417", "WCS315", "CHA0"))

df_summary$water_regime <- factor(df_summary$water_regime, c("WW", "DR"))

df_summary$Phylum <- factor(df_summary$Phylum, c("Proteobacteria", "Planctomycetota", "Actinobacteriota","Bacteroidota", "Verrucomicrobiota","Acidobacteriota", "Chloroflexi", "Patescibacteria", "Gemmatimonadota", "Myxococcota", "Firmicutes", "Bdellovibrionota", "Armatimonadota", "Cyanobacteria", "Dependentiae","Abditibacteriota", "Nitrospirota", "WPS-2", "Hydrogenedentes", "NA"))


ggplot(df_summary, aes(x = n_ASVs, y = Phylum, fill = Treatment)) +
  geom_col(position = position_dodge(width = 0.95)) +
  scale_fill_manual(values = c("#C9DC53", "#F99F58", "#9771A5")) +
  facet_wrap(~water_regime) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    x = "Number of ASVs (Enriched + / Depleted -)",
    y = "Phylum",
    fill = "Treatment"
  ) +
  theme_bw() + 
  theme(axis.text.y = element_text(colour = "black", size = 12, face = "bold"))+
  theme(axis.text.x = element_text(colour = "black", size = 12, angle = 0, hjust = 1, face = "plain"))


mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./5_Brassica_oleracea_exp/results/plots/diff_abundance",
  scale = 1.8,
  width = 1800,
  height = 1000,
  units = "px",
  dpi = 300),
x=c("Fig.5c_n_ASVs.svg","Fig.5c_n_ASVs.png"))



```

#Plotting LFC of ASVs enriched/depleated at Phyllum level

```{r}
#now plot

df_combined_f$Treatment <- factor(df_combined_f$Treatment, c("Control", "WCS417", "WCS315", "CHA0"))

df_combined_f$water_regime <- factor(df_combined_f$water_regime, c("WW", "DR"))

ggplot(df_combined_f,
       aes(x = Family, y = lfc, fill = Treatment)) +
   geom_point(aes(fill=Treatment), shape=21, color="black", stroke = 0.3,
             size=3, alpha=0.7)+ # individual sample values
  labs(y = "LFC", color = "Treatment") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  scale_fill_manual(values = c("#DC267F", "#FE6100", "#FFB000")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size=0.4)+
  facet_grid(water_regime~Phylum, scales = "free_x", space = "free_x")+
   theme_bw() + 
  theme(axis.text.y = element_text(colour = "black", size = 8, face = "bold"))+
  theme(axis.text.x = element_text(colour = "black", size = 8, angle = 90, hjust = 1, face = "plain"))


#and if I order the phyllum based on more to less represented?

unique(df_combined_f$Phylum) #20 phylums

sum(df_combined_f$Phylum == "Proteobacteria", na.rm = TRUE) #555
sum(df_combined_f$Phylum == "Bacteroidota", na.rm = TRUE) #114
sum(df_combined_f$Phylum == "Patescibacteria", na.rm = TRUE) #51
sum(df_combined_f$Phylum == "Acidobacteriota", na.rm = TRUE) #72
sum(df_combined_f$Phylum == "Verrucomicrobiota", na.rm = TRUE) #114
sum(df_combined_f$Phylum == "Myxococcota", na.rm = TRUE) #33
sum(df_combined_f$Phylum == "Planctomycetota", na.rm = TRUE) #180
sum(df_combined_f$Phylum == "Armatimonadota", na.rm = TRUE) #9
sum(df_combined_f$Phylum == "WPS-2", na.rm = TRUE) #6
sum(df_combined_f$Phylum == "Gemmatimonadota", na.rm = TRUE) #42
sum(df_combined_f$Phylum == "Hydrogenedentes", na.rm = TRUE) #3
sum(df_combined_f$Phylum == "Nitrospirota", na.rm = TRUE) #6
sum(df_combined_f$Phylum == "Bdellovibrionota", na.rm = TRUE) #18
sum(df_combined_f$Phylum == "Actinobacteriota", na.rm = TRUE) #129
sum(df_combined_f$Phylum == "Chloroflexi", na.rm = TRUE) #60
sum(df_combined_f$Phylum == "Dependentiae", na.rm = TRUE) #9
sum(df_combined_f$Phylum == "NA", na.rm = TRUE) #0
sum(df_combined_f$Phylum == "Cyanobacteria", na.rm = TRUE) #9
sum(df_combined_f$Phylum == "Firmicutes", na.rm = TRUE) #21
sum(df_combined_f$Phylum == "Abditibacteriota", na.rm = TRUE) #6

#Plot againn
df_combined_f$Treatment <- factor(df_combined_f$Treatment, c("WCS417", "WCS315", "CHA0"))

df_combined_f$water_regime <- factor(df_combined_f$water_regime, c("WW", "DR"))

df_combined_f$Phylum <- factor(df_combined_f$Phylum, c("Proteobacteria", "Planctomycetota", "Actinobacteriota","Bacteroidota", "Verrucomicrobiota","Acidobacteriota", "Chloroflexi", "Patescibacteria", "Gemmatimonadota", "Myxococcota", "Firmicutes", "Bdellovibrionota", "Armatimonadota", "Cyanobacteria", "Dependentiae","Abditibacteriota", "Nitrospirota", "WPS-2", "Hydrogenedentes", "NA"))

ggplot(df_combined_f,
       aes(x = Family, y = lfc, fill = Treatment)) +
   geom_point(aes(fill=Treatment), shape=21, color="black", stroke = 0.3,
             size=3, alpha=0.7)+ # individual sample values
  labs(y = "LFC", color = "Treatment") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  scale_fill_manual(values = c("#C9DC53", "#F99F58", "#9771A5")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size=0.4)+
  facet_grid(water_regime~Phylum, scales = "free_x", space = "free_x")+
   theme_bw() + 
  theme(axis.text.y = element_text(colour = "black", size = 8, face = "bold"))+
  theme(axis.text.x = element_text(colour = "black", size = 8, angle = 90, hjust = 1, face = "plain"))

mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./5_Brassica_oleracea_exp/results/plots/diff_abundance",
  scale = 1.8,
  width = 3500,
  height = 1250,
  units = "px",
  dpi = 300),
x=c("Fig.S6_LFC.svg","Fig.S6_LFC.png"))




```



##data for venn diagrams
```{r}

## WW ##

#UP#
WW_WCS417_up <- ancombc_WW_l$df_sig_asv_l$WCS417 %>% filter(lfc>0) %>%     rownames_to_column("ASV_WCS417")

write.csv(WW_WCS417_up, "WW_WCS417_UP.csv", row.names = FALSE)

WW_WCS315_up <- ancombc_WW_l$df_sig_asv_l$WCS315 %>% filter(lfc>0) %>%     rownames_to_column("ASV_WCS315")

write.csv(WW_WCS315_up, "WW_WCS315_UP.csv", row.names = FALSE)

WW_CHA0_up <- ancombc_WW_l$df_sig_asv_l$CHA0 %>% filter(lfc>0)  %>%     rownames_to_column("ASV_CHA0")

write.csv(WW_CHA0_up, "WW_CHA0_UP.csv", row.names = FALSE)

#DOWN#

WW_WCS417_down <- ancombc_WW_l$df_sig_asv_l$WCS417 %>% filter(lfc<0) %>%     rownames_to_column("ASV_WCS417")

write.csv(WW_WCS417_down, "WW_WCS417_DOWN.csv", row.names = FALSE)

WW_WCS315_down <- ancombc_WW_l$df_sig_asv_l$WCS315 %>% filter(lfc<0) %>%     rownames_to_column("ASV_WCS315")

write.csv(WW_WCS315_down, "WW_WCS315_DOWN.csv", row.names = FALSE)

WW_CHA0_down <- ancombc_WW_l$df_sig_asv_l$CHA0 %>% filter(lfc<0) %>%     rownames_to_column("ASV_CHA0")

write.csv(WW_CHA0_down, "WW_CHA0_DOWN.csv", row.names = FALSE)



## DR ##

#UP#
DR_WCS417_up <- ancombc_DR_l$df_sig_asv_l$WCS417 %>% filter(lfc>0) %>%     rownames_to_column("ASV_WCS417")

write.csv(DR_WCS417_up, "DR_WCS417_UP.csv", row.names = FALSE)

DR_WCS315_up <- ancombc_DR_l$df_sig_asv_l$WCS315 %>% filter(lfc>0) %>%     rownames_to_column("ASV_WCS315")

write.csv(DR_WCS315_up, "DR_WCS315_UP.csv", row.names = FALSE)

DR_CHA0_up <- ancombc_DR_l$df_sig_asv_l$CHA0 %>% filter(lfc>0)  %>%     rownames_to_column("ASV_CHA0")

write.csv(DR_CHA0_up, "DR_CHA0_UP.csv", row.names = FALSE)

#DOWN#

DR_WCS417_down <- ancombc_DR_l$df_sig_asv_l$WCS417 %>% filter(lfc<0) %>%     rownames_to_column("ASV_WCS417")

write.csv(DR_WCS417_down, "DR_WCS417_DOWN.csv", row.names = FALSE)

DR_WCS315_down <- ancombc_DR_l$df_sig_asv_l$WCS315 %>% filter(lfc<0) %>%     rownames_to_column("ASV_WCS315")

write.csv(DR_WCS315_down, "DR_WCS315_DOWN.csv", row.names = FALSE)

DR_CHA0_down <- ancombc_DR_l$df_sig_asv_l$CHA0 %>% filter(lfc<0) %>%     rownames_to_column("ASV_CHA0")

write.csv(DR_CHA0_down, "DR_CHA0_DOWN.csv", row.names = FALSE)



```


