---
title: "04_differential_abundance_ANCOMBC"
author: "Zulema Carracedo Lorenzo"
date: "2025-03-12"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Load packages
```{r}
library(phyloseq)
library(ggplot2)
library(dplyr)
library(tidyr)
library(vegan)
library(UpSetR)
library(purrr)
library(tibble)
library(tidyr)
library(usefun)
library(writexl)

#Differential abundance
library(metagMisc) #for CSS transformation & prevalence filter
library(metagenomeSeq)
library(DESeq2) #differential abundance
library(zinbwave) #zero-inflated
library(ANCOMBC) #ANCOM
#library(GUniFrac) #ZicoSeq & rarefaction PROBLEM
library(ALDEx2) #Aldex2
library(Maaslin2) #MaAslin2
#library(corncob) #not able to install

#Heatmaps
library(ComplexHeatmap)

```

#Load phyloseq object
```{r}
#For differential abundance analysis we use an unnormalize phyloseq object

load("./results/phyloseq/ps_filtered.RData") 

#We will only use the samples from the bulk soil

ps_filtered_bulk <-  subset_samples(ps_filtered, Sample_type %in% "bulk_soil")
ps_filtered_bulk <- prune_taxa(taxa_sums(ps_filtered_bulk) >0, ps_filtered_bulk)

rm(ps_filtered)
```

#Define pairwise comparisons
```{r}

#We need to divide our data into WW and DR

ps_WW_bulk <- subset_samples(ps_filtered_bulk, water_regime %in% "WW")
ps_WW_bulk <- prune_taxa(taxa_sums(ps_WW_bulk) >0, ps_WW_bulk)

ps_DR_bulk <- subset_samples(ps_filtered_bulk, water_regime %in% "DR")
ps_DR_bulk <- prune_taxa(taxa_sums(ps_DR_bulk) >0, ps_DR_bulk)


```



# Filtering low prevalence taxa

```{r}
ps_filtered_bulk #48,027 ASV

### WW ###
ps_WW_bulk #34,464 ASVs

#First, check how many zero's
zero_percentage <- function(x){
   df <- data.frame(otu_table(x))
   zero_percentage <- (sum(colSums(df == 0))) / (nrow(df) * ncol(df))
   return(zero_percentage)
}

zero_percentage(ps_WW_bulk) #84.31%

# Now, drop taxa present in at least 20% of the samples
ps_WW_bulk_f <- metagMisc::phyloseq_filter_prevalence(ps_WW_bulk, prev.trh = 0.20, #at least 20% of samples
                                          abund.trh = NULL) #no filtering by abundance 

#check
ps_WW_bulk_f #6,846 ASVs

zero_percentage(ps_WW_bulk_f) # 43.75%


### DR ###
ps_DR_bulk #29,367 ASVs

#First, check how many zero's
zero_percentage(ps_DR_bulk) #83.15%

# Now, drop taxa present in at least 20% of the samples
ps_DR_bulk_f <- metagMisc::phyloseq_filter_prevalence(ps_DR_bulk, prev.trh = 0.20, #at least 20% of samples
                                          abund.trh = NULL) #no filtering by abundance 

#check
ps_DR_bulk_f #6,122 ASVs

zero_percentage(ps_DR_bulk_f) # 41.34%

rm(ps_WW_bulk,ps_DR_bulk)


```


# 1. ANCOM-BC

ANCOM-BC (Analysis of Compositions of Microbiomes with Bias Correction) is a method that takes into account the compositionality of microbiome data by using an additive log-ratio transformation. A ratio is calculated for each taxon with a reference taxon with low variance present across all samples. The differences between the ratios are compared and tested statistically to determine whether a taxon is DA between two conditions or not. It is considered a stringent method. Input is non-rarefied data. 

This is a good explanation of ANCOM-BC : https://github.com/FrederickHuangLin/ANCOM-BC-Code-Archive

##WW
```{r}

# Note that by default, levels of a categorical variable in R are sorted alphabetically. In this case, the reference level for `treatment` will be`CHA0`. To manually change the reference level, use:

ps_WW_bulk_f@sam_data$treatment = factor(ps_WW_bulk_f@sam_data$treatment, levels = c("Control", "WCS417", "WCS315", "CHA0"))

# You can verify the change by checking:
levels(ps_WW_bulk_f@sam_data$treatment)

#Calculate ANCOm-BC

  
  set.seed(123)
  ancom <- ancombc(phyloseq = ps_WW_bulk_f, 
                      formula = "treatment", 
                      p_adj_method = "fdr", #same as "BH" Benjamini & Hochberg
                      zero_cut = 0.9, #we are removing ASVs that have more than 90% of 0s
                      lib_cut = 0, 
                      group = "treatment", 
                      struc_zero = TRUE, #with structural zero's
                      neg_lb = FALSE, #classifies structural zero's if it's zero and not its asymptotic lower bound 
                      tol = 1e-5,               
                      max_iter = 100, 
                      conserve = TRUE, 
                      alpha = 0.05)

  save(ancom, file="./results/r_objects/differential_abundance/ancom_bc_bulk_WW.RData")
  load(file="./results/r_objects/differential_abundance/ancom_bc_bulk_WW.RData")
  
  #get results from ANCOM-BC into df
 df_ancom_WW <- data.frame(
  ASV = row.names(ancom$res$beta), #get ASV names
  lfc = unlist(ancom$res$beta),
  se = unlist(ancom$res$se),
  W = unlist(ancom$res$W),
  p_val = unlist(ancom$res$p_val),
  q_val = unlist(ancom$res$q_val),
  diff_abn = unlist(ancom$res$diff_abn))
 
 #Row names to column and fix teatment names
 
 df_ancom_WW <- rownames_to_column(df_ancom_WW)
 
 df_ancom_WW <- df_ancom_WW %>% rename("rowname" = "treatment")
 
df_ancom_WW <- df_ancom_WW %>%
   mutate(Treatment =
            case_when(grepl("WCS417", df_ancom_WW$treatment) ~  "WCS417",
                      grepl("WCS315", df_ancom_WW$treatment) ~  "WCS315",
                     grepl("CHA0", df_ancom_WW$treatment) ~  "CHA0" 
                      ))

df_ancom_WW <- df_ancom_WW[,2:9] #removes wrong treatment column

#store in a list a singel dataframe per treatment

df_ancom_WW_417 <- df_ancom_WW %>% filter(Treatment == "WCS417")
df_ancom_WW_315 <- df_ancom_WW %>% filter(Treatment == "WCS315")
df_ancom_WW_CHA0 <- df_ancom_WW %>% filter(Treatment == "CHA0")

df_ancom_WW_bulk_l <- list(df_ancom_WW_417, df_ancom_WW_315, df_ancom_WW_CHA0)
names(df_ancom_WW_bulk_l) <- c("WCS417", "WCS315", "CHA0")

rm(df_ancom_WW_417, df_ancom_WW_315, df_ancom_WW_CHA0) 

#place ASVs as rownames, for some reason didn't allow me to do it with lapply

  df_ancom_WW_bulk_l$WCS417 <- df_ancom_WW_bulk_l$WCS417 %>% 
              remove_rownames()%>%
              column_to_rownames(var="ASV")

  df_ancom_WW_bulk_l$WCS315 <- df_ancom_WW_bulk_l$WCS315 %>% 
              remove_rownames()%>%
              column_to_rownames(var="ASV")
  
    df_ancom_WW_bulk_l$CHA0 <- df_ancom_WW_bulk_l$CHA0 %>% 
              remove_rownames()%>%
              column_to_rownames(var="ASV")
    
  
  #join df's from ANCOMBC and phyloseq
    
  df_ancom_WW_bulk_l$WCS417  <- cbind(as(df_ancom_WW_bulk_l$WCS417, "data.frame"),
                  as(tax_table(ps_WW_bulk_f)[rownames(df_ancom_WW_bulk_l$WCS417), ], "matrix"))
  
  df_ancom_WW_bulk_l$WCS315  <- cbind(as(df_ancom_WW_bulk_l$WCS315, "data.frame"),
                  as(tax_table(ps_WW_bulk_f)[rownames(df_ancom_WW_bulk_l$WCS315), ], "matrix"))
  
  df_ancom_WW_bulk_l$CHA0  <- cbind(as(df_ancom_WW_bulk_l$CHA0, "data.frame"),
                  as(tax_table(ps_WW_bulk_f)[rownames(df_ancom_WW_bulk_l$CHA0), ], "matrix"))

  
  #keep only those with adjusted p value of <0.05 ("q_val") 

  
  df_ancom_WW_ps_sig_WCS417 <- df_ancom_WW_bulk_l$WCS417 %>% 
   dplyr::filter(q_val < 0.05)%>%
   mutate(test="3_ancombc") #1045
  
    df_ancom_WW_ps_sig_WCS315 <- df_ancom_WW_bulk_l$WCS315 %>% 
   dplyr::filter(q_val < 0.05)%>%
   mutate(test="3_ancombc") #627
    
      df_ancom_WW_ps_sig_CHA0 <- df_ancom_WW_bulk_l$CHA0 %>% 
   dplyr::filter(q_val < 0.05)%>%
   mutate(test="3_ancombc") #629
  
df_ancom_WW_ps_sig_l <- list(df_ancom_WW_ps_sig_WCS417, df_ancom_WW_ps_sig_WCS315, df_ancom_WW_ps_sig_CHA0) #list of lists of significant ASVs detectec by ANCOM by each treatment

names(df_ancom_WW_ps_sig_l) <- c("WCS417", "WCS315", "CHA0")



#
#check how many DA asvs were

sum(df_ancom_WW_bulk_l$WCS417$diff_abn == "TRUE") #1,045
sum(df_ancom_WW_bulk_l$WCS315$diff_abn == "TRUE") #627
sum(df_ancom_WW_bulk_l$CHA0$diff_abn == "TRUE") #629

#Save test results and significan ASV list in a single object

ancombc_WW_l <- list(results= df_ancom_WW_bulk_l, df_sig_asv_l = df_ancom_WW_ps_sig_l)

#check which ones 
lapply(ancom_bc_l, function(x) rownames(x$df_sig_asv)) 

##Save
#save(ancom_bc_l, file="./results/r_objects/differential_abundance/1_ancombc_l.RData")
#load(file="./results/r_objects/differential_abundance/1_ancombc_l.RData")

#save list of ASVs per strains
ancombc_WW_WCS417 <- ancombc_WW_l$df_sig_asv_l$WCS417
write_xlsx(ancombc_WW_WCS417, "./results/tables/diff_abundance/ancombc_WW_WCS417.xlsx")

ancombc_WW_WCS315 <- ancombc_WW_l$df_sig_asv_l$WCS315
write_xlsx(ancombc_WW_WCS315, "./results/tables/diff_abundance/ancombc_WW_WCS315.xlsx")

ancombc_WW_CHA0 <- ancombc_WW_l$df_sig_asv_l$CHA0
write_xlsx(ancombc_WW_CHA0, "./results/tables/diff_abundance/ancombc_WW_CHA0.xlsx")

         
```

##DR
```{r}

# Note that by default, levels of a categorical variable in R are sorted alphabetically. In this case, the reference level for `treatment` will be`CHA0`. To manually change the reference level, use:

ps_DR_bulk_f@sam_data$treatment = factor(ps_DR_bulk_f@sam_data$treatment, levels = c("Control", "WCS417", "WCS315", "CHA0"))

# You can verify the change by checking:
levels(ps_DR_bulk_f@sam_data$treatment)

#Calculate ANCOm-BC

  
  set.seed(123)
  ancom <- ancombc(phyloseq = ps_DR_bulk_f, 
                      formula = "treatment", 
                      p_adj_method = "fdr", #same as "BH" Benjamini & Hochberg
                      zero_cut = 0.9, #we are removing ASVs that have more than 90% of 0s
                      lib_cut = 0, 
                      group = "treatment", 
                      struc_zero = TRUE, #with structural zero's
                      neg_lb = FALSE, #classifies structural zero's if it's zero and not its asymptotic lower bound 
                      tol = 1e-5,               
                      max_iter = 100, 
                      conserve = TRUE, 
                      alpha = 0.05)

  save(ancom, file="./results/r_objects/differential_abundance/ancom_bc_bulk_DR.RData")
  load(file="./results/r_objects/differential_abundance/ancom_bc_bulk_DR.RData")
  
  #get results from ANCOM-BC into df
 df_ancom_DR <- data.frame(
  ASV = row.names(ancom$res$beta), #get ASV names
  lfc = unlist(ancom$res$beta),
  se = unlist(ancom$res$se),
  W = unlist(ancom$res$W),
  p_val = unlist(ancom$res$p_val),
  q_val = unlist(ancom$res$q_val),
  diff_abn = unlist(ancom$res$diff_abn))
 
 #Row names to column and fix teatment names
 
 df_ancom_DR <- rownames_to_column(df_ancom_DR)
 
 df_ancom_DR <- df_ancom_DR %>% rename("rowname" = "treatment")
 
df_ancom_DR <- df_ancom_DR %>%
   mutate(Treatment =
            case_when(grepl("WCS417", df_ancom_DR$treatment) ~  "WCS417",
                      grepl("WCS315", df_ancom_DR$treatment) ~  "WCS315",
                     grepl("CHA0", df_ancom_DR$treatment) ~  "CHA0" 
                      ))

df_ancom_DR <- df_ancom_DR[,2:9] #removes wrong treatment column

#store in a list a singel dataframe per treatment

df_ancom_DR_417 <- df_ancom_DR %>% filter(Treatment == "WCS417")
df_ancom_DR_315 <- df_ancom_DR %>% filter(Treatment == "WCS315")
df_ancom_DR_CHA0 <- df_ancom_DR %>% filter(Treatment == "CHA0")

df_ancom_DR_bulk_l <- list(df_ancom_DR_417, df_ancom_DR_315, df_ancom_DR_CHA0)
names(df_ancom_DR_bulk_l) <- c("WCS417", "WCS315", "CHA0")

rm(df_ancom_DR_417, df_ancom_DR_315, df_ancom_DR_CHA0) 

#place ASVs as rownames, for some reason didn't allow me to do it with lapply

  df_ancom_DR_bulk_l$WCS417 <- df_ancom_DR_bulk_l$WCS417 %>% 
              remove_rownames()%>%
              column_to_rownames(var="ASV")

  df_ancom_DR_bulk_l$WCS315 <- df_ancom_DR_bulk_l$WCS315 %>% 
              remove_rownames()%>%
              column_to_rownames(var="ASV")
  
    df_ancom_DR_bulk_l$CHA0 <- df_ancom_DR_bulk_l$CHA0 %>% 
              remove_rownames()%>%
              column_to_rownames(var="ASV")
    
  
  #join df's from ANCOMBC and phyloseq
    
  df_ancom_DR_bulk_l$WCS417  <- cbind(as(df_ancom_DR_bulk_l$WCS417, "data.frame"),
                  as(tax_table(ps_DR_bulk_f)[rownames(df_ancom_DR_bulk_l$WCS417), ], "matrix"))
  
  df_ancom_DR_bulk_l$WCS315  <- cbind(as(df_ancom_DR_bulk_l$WCS315, "data.frame"),
                  as(tax_table(ps_DR_bulk_f)[rownames(df_ancom_DR_bulk_l$WCS315), ], "matrix"))
  
  df_ancom_DR_bulk_l$CHA0  <- cbind(as(df_ancom_DR_bulk_l$CHA0, "data.frame"),
                  as(tax_table(ps_DR_bulk_f)[rownames(df_ancom_DR_bulk_l$CHA0), ], "matrix"))

  
  #keep only those with adjusted p value of <0.05 ("q_val") 

  
  df_ancom_DR_ps_sig_WCS417 <- df_ancom_DR_bulk_l$WCS417 %>% 
   dplyr::filter(q_val < 0.05)%>%
   mutate(test="3_ancombc") #695
  
    df_ancom_DR_ps_sig_WCS315 <- df_ancom_DR_bulk_l$WCS315 %>% 
   dplyr::filter(q_val < 0.05)%>%
   mutate(test="3_ancombc") #630
    
      df_ancom_DR_ps_sig_CHA0 <- df_ancom_DR_bulk_l$CHA0 %>% 
   dplyr::filter(q_val < 0.05)%>%
   mutate(test="3_ancombc") #386
  
df_ancom_DR_ps_sig_l <- list(df_ancom_DR_ps_sig_WCS417, df_ancom_DR_ps_sig_WCS315, df_ancom_DR_ps_sig_CHA0) #list of lists of significant ASVs detectec by ANCOM by each treatment

names(df_ancom_DR_ps_sig_l) <- c("WCS417", "WCS315", "CHA0")


#check how many DA asvs were

sum(df_ancom_DR_bulk_l$WCS417$diff_abn == "TRUE") #695
sum(df_ancom_DR_bulk_l$WCS315$diff_abn == "TRUE") #630
sum(df_ancom_DR_bulk_l$CHA0$diff_abn == "TRUE") #386

#Save test results and significan ASV list in a single object

ancombc_DR_l <- list(results= df_ancom_DR_bulk_l, df_sig_asv_l = df_ancom_DR_ps_sig_l)

#check which ones 
#lapply(ancom_bc_l, function(x) rownames(x$df_sig_asv)) 

#Save
#save(ancom_bc_l, file="./results/r_objects/differential_abundance/1_ancombc_l.RData")
#load(file="./results/r_objects/differential_abundance/1_ancombc_l.RData")

#save list of ASVs per strains
ancombc_DR_WCS417 <- ancombc_DR_l$df_sig_asv_l$WCS417
write_xlsx(ancombc_DR_WCS417, "./results/tables/diff_abundance/ancombc_DR_WCS417.xlsx")

ancombc_DR_WCS315 <- ancombc_DR_l$df_sig_asv_l$WCS315
write_xlsx(ancombc_DR_WCS315, "./results/tables/diff_abundance/ancombc_DR_WCS315.xlsx")

ancombc_DR_CHA0 <- ancombc_DR_l$df_sig_asv_l$CHA0
write_xlsx(ancombc_DR_CHA0, "./results/tables/diff_abundance/ancombc_DR_CHA0.xlsx")

         
```


#Plotting LFC

##WW

```{r}

    ancombc_WW_l

WCS417_df <- ancombc_WW_l$df_sig_asv_l$WCS417 %>% 
  rownames_to_column(var = "ASV")

WCS315_df <- ancombc_WW_l$df_sig_asv_l$WCS315 %>% 
  rownames_to_column(var = "ASV")

CHA0_df <- ancombc_WW_l$df_sig_asv_l$CHA0 %>% 
  rownames_to_column(var = "ASV")

#join in a single dataframe
df_combined <- rbind(WCS417_df, WCS315_df, CHA0_df)

# Summarize the data to get one value per family+treatment
df_summarized <- df_combined %>%
  group_by(Family,Treatment, Phylum) %>%
  summarize(lfc = mean(lfc, na.rm = TRUE), .groups = "drop")  # Change mean() if needed
  
#plot

ggplot(df_summarized, aes(x = Order, y = lfc, color = Treatment)) +
  geom_point(size = 3, alpha = 0.8, position = position_dodge(width = 0.5)) +  # Dodge to separate treatments
  facet_grid(Order ~ Phylum) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Order", y = "Log Fold Change (LFC)", color = "Treatment") +
  scale_color_brewer(palette = "Set1")



ggplot(df_summarized,
       aes(x = Family, y = lfc, fill = Treatment)) +
   geom_point(aes(fill=Treatment), shape=21, color="black", stroke = 0.3,
             size=3, alpha=0.7)+ # individual sample values
  labs(y = "LFC", color = "Treatment") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  #scale_color_manual(values=color_JAtreatments)+
  #scale_fill_manual(values=color_JAtreatments)+
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size=0.4)+
  facet_grid(~Phylum, scales = "free_x", space = "free_x")+
  #axis_looks+
  theme(axis.text.y = element_text(colour = "black", size = 8, face = "bold"))+
  theme(axis.text.x = element_text(colour = "black", size = 8, angle = 90, hjust = 1, face = "plain"))

mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./results/plots/differential_abundance",
  scale = 1.8,
  width = 4000,
  height = 1000,
  units = "px",
  dpi = 300),
x=c("lfc_bulk_WW_March25.svg","lfc_bulk_WW_March25.png"))


# What if I select only the ASVs that are significant for the 3 treatments

df_combined

#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- WCS417_df$ASV #1045 ASVs
DA_WCS315_vector <- WCS315_df$ASV #627 ASVs
DA_CHA0_vector <- CHA0_df$ASV #629 ASVs

#I want to keep only the ASVs  are differentially abundance under the three treatments when compared to control, so keep only ASVs that are present in the 3 vectors

DA_all_treat_vector <- Reduce(intersect, list(DA_WCS417_vector,DA_WCS315_vector,DA_CHA0_vector)) #261 ASVs

df_combined_f <- df_combined %>%
  filter(ASV %in% DA_all_treat_vector)

#now plot

df_combined_f$Treatment <- factor(df_combined_f$Treatment, c("Control", "WCS417", "WCS315", "CHA0"))

ggplot(df_combined_f,
       aes(x = Family, y = lfc, fill = Treatment)) +
   geom_point(aes(fill=Treatment), shape=21, color="black", stroke = 0.3,
             size=3, alpha=0.7)+ # individual sample values
  labs(y = "LFC", color = "Treatment") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  scale_fill_manual(values = c("#DC267F", "#FE6100", "#FFB000")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size=0.4)+
  facet_grid(~Phylum, scales = "free_x", space = "free_x")+
   theme_bw() + 
  theme(axis.text.y = element_text(colour = "black", size = 8, face = "bold"))+
  theme(axis.text.x = element_text(colour = "black", size = 8, angle = 90, hjust = 1, face = "plain"))

mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./results/plots/differential_abundance",
  scale = 1.8,
  width = 4000,
  height = 1000,
  units = "px",
  dpi = 300),
x=c("lfc_bulk_WW_shared_March25.svg","lfc_bulk_WW_shared_March25.png"))

```

##DR

```{r}

    ancombc_DR_l

WCS417_df <- ancombc_DR_l$df_sig_asv_l$WCS417 %>% 
  rownames_to_column(var = "ASV")

WCS315_df <- ancombc_DR_l$df_sig_asv_l$WCS315 %>% 
  rownames_to_column(var = "ASV")

CHA0_df <- ancombc_DR_l$df_sig_asv_l$CHA0 %>% 
  rownames_to_column(var = "ASV")

#join in a single dataframe
df_combined_DR <- rbind(WCS417_df, WCS315_df, CHA0_df)

# Summarize the data to get one value per family+treatment
df_summarized_DR <- df_combined_DR %>%
  group_by(Family,Treatment, Phylum) %>%
  summarize(lfc = mean(lfc, na.rm = TRUE), .groups = "drop")  # Change mean() if needed
  
#plot

ggplot(df_summarized_DR,
       aes(x = Family, y = lfc, fill = Treatment)) +
   geom_point(aes(fill=Treatment), shape=21, color="black", stroke = 0.3,
             size=3, alpha=0.7)+ # individual sample values
  labs(y = "LFC", color = "Treatment") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  #scale_color_manual(values=color_JAtreatments)+
  #scale_fill_manual(values=color_JAtreatments)+
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size=0.4)+
  facet_grid(~Phylum, scales = "free_x", space = "free_x")+
  #axis_looks+
  theme(axis.text.y = element_text(colour = "black", size = 8, face = "bold"))+
  theme(axis.text.x = element_text(colour = "black", size = 8, angle = 90, hjust = 1, face = "plain"))

mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./results/plots/differential_abundance",
  scale = 1.8,
  width = 4000,
  height = 1000,
  units = "px",
  dpi = 300),
x=c("lfc_bulk_DR_March25.svg","lfc_bulk_DR_March25.png"))


# What if I select only the ASVs that are significant for the 3 treatments

df_combined_DR

#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- WCS417_df$ASV #695 ASVs
DA_WCS315_vector <- WCS315_df$ASV #630 ASVs
DA_CHA0_vector <- CHA0_df$ASV #386 ASVs

#I want to keep only the ASVs  are differentially abundance under the three treatments when compared to control, so keep only ASVs that are present in the 3 vectors

DA_all_treat_vector <- Reduce(intersect, list(DA_WCS417_vector,DA_WCS315_vector,DA_CHA0_vector)) #386 ASVs

df_combined_DR_f <- df_combined_DR %>%
  filter(ASV %in% DA_all_treat_vector)

#now plot

df_combined_DR_f$Treatment <- factor(df_combined_DR_f$Treatment, c("Control", "WCS417", "WCS315", "CHA0"))

ggplot(df_combined_DR_f,
       aes(x = Family, y = lfc, fill = Treatment)) +
   geom_point(aes(fill=Treatment), shape=21, color="black", stroke = 0.3,
             size=3, alpha=0.7)+ # individual sample values
  labs(y = "LFC", color = "Treatment") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  scale_fill_manual(values = c("#DC267F", "#FE6100", "#FFB000")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size=0.4)+
  facet_grid(~Phylum, scales = "free_x", space = "free_x")+
   theme_bw() + 
  theme(axis.text.y = element_text(colour = "black", size = 8, face = "bold"))+
  theme(axis.text.x = element_text(colour = "black", size = 8, angle = 90, hjust = 1, face = "plain"))

mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./results/plots/differential_abundance",
  scale = 1.8,
  width = 4000,
  height = 1000,
  units = "px",
  dpi = 300),
x=c("lfc_bulk_DR_shared_March25.svg","lfc_bulk_DR_shared_March25.png"))

```

## WW+DR
##Shared among 3 strains
```{r}
#Now let's try to make a plot with both WW and DR data

df_combined_DR_f
df_combined_WW_f <- df_combined_f

#add a column for the water regime

df_combined_DR_f <- df_combined_DR_f %>% mutate(water_regime = "DR")
df_combined_WW_f <- df_combined_WW_f %>% mutate(water_regime = "WW")

#join the df

df_combined_f <- rbind(df_combined_DR_f, df_combined_WW_f)

# Optional — remove prefixes like "p__", "f__", etc., for cleaner plot labels
clean_taxonomy <- function(x) {
  x <- trimws(x)  # Remove leading/trailing spaces
  gsub("^\\w__", "", x)  # Remove the prefix
}

# Apply the function with mutate
df_combined_f <- df_combined_f %>%
  mutate(across(c(Phylum, Family, Genus), clean_taxonomy))


#now plot

df_combined_f$Treatment <- factor(df_combined_f$Treatment, c("Control", "WCS417", "WCS315", "CHA0"))

df_combined_f$water_regime <- factor(df_combined_f$water_regime, c("WW", "DR"))

ggplot(df_combined_f,
       aes(x = Family, y = lfc, fill = Treatment)) +
   geom_point(aes(fill=Treatment), shape=21, color="black", stroke = 0.3,
             size=3, alpha=0.7)+ # individual sample values
  labs(y = "LFC", color = "Treatment") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  scale_fill_manual(values = c("#DC267F", "#FE6100", "#FFB000")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size=0.4)+
  facet_grid(water_regime~Phylum, scales = "free_x", space = "free_x")+
   theme_bw() + 
  theme(axis.text.y = element_text(colour = "black", size = 8, face = "bold"))+
  theme(axis.text.x = element_text(colour = "black", size = 8, angle = 90, hjust = 1, face = "plain"))

mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./results/plots/differential_abundance",
  scale = 1.8,
  width = 4000,
  height = 1000,
  units = "px",
  dpi = 300),
x=c("lfc_bulk_shared_March25.svg","lfc_bulk_shared_March25.png"))

#and if I order the phyllum based on more to less represented?

unique(df_combined_f$Phylum) #20 phylums

sum(df_combined_f$Phylum == "Proteobacteria", na.rm = TRUE) #555
sum(df_combined_f$Phylum == "Bacteroidota", na.rm = TRUE) #114
sum(df_combined_f$Phylum == "Patescibacteria", na.rm = TRUE) #51
sum(df_combined_f$Phylum == "Acidobacteriota", na.rm = TRUE) #72
sum(df_combined_f$Phylum == "Verrucomicrobiota", na.rm = TRUE) #114
sum(df_combined_f$Phylum == "Myxococcota", na.rm = TRUE) #33
sum(df_combined_f$Phylum == "Planctomycetota", na.rm = TRUE) #180
sum(df_combined_f$Phylum == "Armatimonadota", na.rm = TRUE) #9
sum(df_combined_f$Phylum == "WPS-2", na.rm = TRUE) #6
sum(df_combined_f$Phylum == "Gemmatimonadota", na.rm = TRUE) #42
sum(df_combined_f$Phylum == "Hydrogenedentes", na.rm = TRUE) #3
sum(df_combined_f$Phylum == "Nitrospirota", na.rm = TRUE) #6
sum(df_combined_f$Phylum == "Bdellovibrionota", na.rm = TRUE) #18
sum(df_combined_f$Phylum == "Actinobacteriota", na.rm = TRUE) #129
sum(df_combined_f$Phylum == "Chloroflexi", na.rm = TRUE) #60
sum(df_combined_f$Phylum == "Dependentiae", na.rm = TRUE) #9
sum(df_combined_f$Phylum == "NA", na.rm = TRUE) #0
sum(df_combined_f$Phylum == "Cyanobacteria", na.rm = TRUE) #9
sum(df_combined_f$Phylum == "Firmicutes", na.rm = TRUE) #21
sum(df_combined_f$Phylum == "Abditibacteriota", na.rm = TRUE) #6

#Plot againn
df_combined_f$Treatment <- factor(df_combined_f$Treatment, c("WCS417", "WCS315", "CHA0"))

df_combined_f$water_regime <- factor(df_combined_f$water_regime, c("WW", "DR"))

df_combined_f$Phylum <- factor(df_combined_f$Phylum, c("Proteobacteria", "Planctomycetota", "Actinobacteriota","Bacteroidota", "Verrucomicrobiota","Acidobacteriota", "Chloroflexi", "Patescibacteria", "Gemmatimonadota", "Myxococcota", "Firmicutes", "Bdellovibrionota", "Armatimonadota", "Cyanobacteria", "Dependentiae","Abditibacteriota", "Nitrospirota", "WPS-2", "Hydrogenedentes", "NA"))

ggplot(df_combined_f,
       aes(x = Family, y = lfc, fill = Treatment)) +
   geom_point(aes(fill=Treatment), shape=21, color="black", stroke = 0.3,
             size=3, alpha=0.7)+ # individual sample values
  labs(y = "LFC", color = "Treatment") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  scale_fill_manual(values = c("#DC267F", "#FE6100", "#FFB000")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size=0.4)+
  facet_grid(water_regime~Phylum, scales = "free_x", space = "free_x")+
   theme_bw() + 
  theme(axis.text.y = element_text(colour = "black", size = 8, face = "bold"))+
  theme(axis.text.x = element_text(colour = "black", size = 8, angle = 90, hjust = 1, face = "plain"))

mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./results/plots/differential_abundance",
  scale = 1.8,
  width = 3500,
  height = 1250,
  units = "px",
  dpi = 300),
x=c("lfc_bulk_FINAL_March25.svg","lfc_bulk_FINAL_March25.png"))


#make a simplify version for the main text

#Option 1
# Sum LCF values per Phylum and Treatment
summed_df <- df_combined_f %>%
  group_by(Treatment, Phylum, water_regime) %>%
  summarise(LCF_sum = sum(lfc), .groups = "drop")

# Plot
summed_df$Treatment <- factor(summed_df$Treatment, c("WCS417", "WCS315", "CHA0"))

summed_df$water_regime <- factor(summed_df$water_regime, c("WW", "DR"))

summed_df$Phylum <- factor(summed_df$Phylum, c("Proteobacteria", "Planctomycetota", "Actinobacteriota","Bacteroidota", "Verrucomicrobiota","Acidobacteriota", "Chloroflexi", "Patescibacteria", "Gemmatimonadota", "Myxococcota", "Firmicutes", "Bdellovibrionota", "Armatimonadota", "Cyanobacteria", "Dependentiae","Abditibacteriota", "Nitrospirota", "WPS-2", "Hydrogenedentes", "NA"))

ggplot(summed_df, aes(x = LCF_sum, y = Phylum, fill = Treatment)) +
  geom_col(size = 3, position = position_dodge(width = 0.95)) +
  facet_wrap(~ water_regime) +
   scale_fill_manual(values = c("#C9DC53", "#F99F58", "#9771A5")) +
  theme_minimal() +
  labs(
    x = "Cumulative LCF",
    y = "Phylum",
    fill = "Treatment"
  ) +
  theme_bw() + 
  theme(axis.text.y = element_text(colour = "black", size = 12, face = "bold"))+
  theme(axis.text.x = element_text(colour = "black", size = 12, angle = 0, hjust = 1, face = "plain"))

mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./results/plots/differential_abundance",
  scale = 1.8,
  width = 1800,
  height = 1000,
  units = "px",
  dpi = 300),
x=c("cumulative_lfc_bulk_May25.svg","cumulative_lfc_bulk_May25.png"))



#Option 2
#check how many ASVs are depleated/enriched per phylun for each treatment and water regime

# Step 1: Add 'status' column to classify LFC direction
df_summary <- df_combined_f %>%
  mutate(status = ifelse(lfc > 0, "Enriched", "Depleted")) %>%
  group_by(Phylum, Treatment, water_regime, status) %>%
  summarise(n_ASVs = n_distinct(ASV), .groups = "drop") %>%
  # Assign negative sign to depleted counts
  mutate(n_ASVs = ifelse(status == "Depleted", -n_ASVs, n_ASVs))

#plot

df_summary$Treatment <- factor(df_summary$Treatment, c("WCS417", "WCS315", "CHA0"))

df_summary$water_regime <- factor(df_summary$water_regime, c("WW", "DR"))

df_summary$Phylum <- factor(df_summary$Phylum, c("Proteobacteria", "Planctomycetota", "Actinobacteriota","Bacteroidota", "Verrucomicrobiota","Acidobacteriota", "Chloroflexi", "Patescibacteria", "Gemmatimonadota", "Myxococcota", "Firmicutes", "Bdellovibrionota", "Armatimonadota", "Cyanobacteria", "Dependentiae","Abditibacteriota", "Nitrospirota", "WPS-2", "Hydrogenedentes", "NA"))


ggplot(df_summary, aes(x = n_ASVs, y = Phylum, fill = Treatment)) +
  geom_col(position = position_dodge(width = 0.95)) +
  scale_fill_manual(values = c("#C9DC53", "#F99F58", "#9771A5")) +
  facet_wrap(~water_regime) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    x = "Number of ASVs (Enriched + / Depleted -)",
    y = "Phylum",
    fill = "Treatment"
  ) +
  theme_bw() + 
  theme(axis.text.y = element_text(colour = "black", size = 12, face = "bold"))+
  theme(axis.text.x = element_text(colour = "black", size = 12, angle = 0, hjust = 1, face = "plain"))


mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./results/plots/differential_abundance",
  scale = 1.8,
  width = 1800,
  height = 1000,
  units = "px",
  dpi = 300),
x=c("n_ASVs_May25.svg","n_ASVs_May255.png"))
```

#check numbers ASVs for all samples, without checking only common ones
```{r}

#WW

  ancombc_WW_l

WCS417_df <- ancombc_WW_l$df_sig_asv_l$WCS417 %>% 
  rownames_to_column(var = "ASV")

WCS315_df <- ancombc_WW_l$df_sig_asv_l$WCS315 %>% 
  rownames_to_column(var = "ASV")

CHA0_df <- ancombc_WW_l$df_sig_asv_l$CHA0 %>% 
  rownames_to_column(var = "ASV")

#join in a single dataframe
df_combined <- rbind(WCS417_df, WCS315_df, CHA0_df)

# Summarize the data to get one value per family+treatment
df_summarized <- df_combined %>%
  group_by(Family,Treatment, Phylum,ASV) %>%
  summarize(lfc = mean(lfc, na.rm = TRUE), .groups = "drop")  # Change mean() if needed

#DR

  ancombc_DR_l

WCS417_df <- ancombc_DR_l$df_sig_asv_l$WCS417 %>% 
  rownames_to_column(var = "ASV")

WCS315_df <- ancombc_DR_l$df_sig_asv_l$WCS315 %>% 
  rownames_to_column(var = "ASV")

CHA0_df <- ancombc_DR_l$df_sig_asv_l$CHA0 %>% 
  rownames_to_column(var = "ASV")

#join in a single dataframe
df_combined_DR <- rbind(WCS417_df, WCS315_df, CHA0_df)

# Summarize the data to get one value per family+treatment
df_summarized_DR <- df_combined_DR %>%
  group_by(Family,Treatment, Phylum,ASV) %>%
  summarize(lfc = mean(lfc, na.rm = TRUE), .groups = "drop")  # Change mean() if needed
    

###
df_summarized #WW
df_summarized_DR #DR

#add a column for the water regime

df_summarized_DR <- df_summarized_DR %>% mutate(water_regime = "DR")
df_summarized_WW <- df_summarized %>% mutate(water_regime = "WW")

#join the df

df_summarized_f <- rbind(df_summarized_DR, df_summarized_WW)

# Optional — remove prefixes like "p__", "f__", etc., for cleaner plot labels
clean_taxonomy <- function(x) {
  x <- trimws(x)  # Remove leading/trailing spaces
  gsub("^\\w__", "", x)  # Remove the prefix
}

#check how many ASVs are depleated/enriched per phylun for each treatment and water regime
#Step 1: Add 'status' column to classify LFC direction
df_summary <- df_summarized_f %>%
  mutate(status = ifelse(lfc > 0, "Enriched", "Depleted")) %>%
  group_by(Phylum, Treatment, water_regime, status) %>%
  summarise(n_ASVs = n_distinct(ASV), .groups = "drop") %>%
  # Assign negative sign to depleted counts
  mutate(n_ASVs = ifelse(status == "Depleted", -n_ASVs, n_ASVs))

# Apply the function with mutate
df_summarized <- df_summary %>%
  mutate(across(Phylum, clean_taxonomy))

df_summary$Treatment <- factor(df_summary$Treatment, c("WCS417", "WCS315", "CHA0"))

df_summary$water_regime <- factor(df_summary$water_regime, c("WW", "DR"))

#df_summary$Phylum <- factor(df_summary$Phylum, c("Proteobacteria", "Planctomycetota", "Actinobacteriota","Bacteroidota", "Verrucomicrobiota","Acidobacteriota", "Chloroflexi", "Patescibacteria", "Gemmatimonadota", "Myxococcota", "Firmicutes", "Bdellovibrionota", "Armatimonadota", "Cyanobacteria", "Dependentiae","Abditibacteriota", "Nitrospirota", "WPS-2", "Hydrogenedentes", "NA"))


ggplot(df_summary, aes(x = n_ASVs, y = Phylum, fill = Treatment)) +
  geom_col(position = position_dodge(width = 0.95)) +
  scale_fill_manual(values = c("#C9DC53", "#F99F58", "#9771A5")) +
  facet_wrap(~water_regime) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    x = "Number of ASVs (Enriched + / Depleted -)",
    y = "Phylum",
    fill = "Treatment"
  ) +
  theme_bw() + 
  theme(axis.text.y = element_text(colour = "black", size = 12, face = "bold"))+
  theme(axis.text.x = element_text(colour = "black", size = 12, angle = 0, hjust = 1, face = "plain"))

```

##Only shared by 2

```{r}
#WW
    ancombc_WW_l

WCS417_df <- ancombc_WW_l$df_sig_asv_l$WCS417 %>% 
  rownames_to_column(var = "ASV")

WCS315_df <- ancombc_WW_l$df_sig_asv_l$WCS315 %>% 
  rownames_to_column(var = "ASV")

CHA0_df <- ancombc_WW_l$df_sig_asv_l$CHA0 %>% 
  rownames_to_column(var = "ASV")

#join in a single dataframe
df_combined <- rbind(WCS417_df, WCS315_df, CHA0_df)


# What if I select only the ASVs that are significant for the 2 treatments

df_combined

#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- WCS417_df$ASV #1045 ASVs
DA_WCS315_vector <- WCS315_df$ASV #627 ASVs
DA_CHA0_vector <- CHA0_df$ASV #629 ASVs

#I want a vector with only the ASVs  are differentially abundance under the three treatments when compared to control, so keep only ASVs that are present in the 3 vectors

DA_all_treat_vector <- Reduce(intersect, list(DA_WCS417_vector,DA_WCS315_vector,DA_CHA0_vector)) #261 ASVs

df_combined_f <- df_combined %>%
  filter(!ASV %in% DA_all_treat_vector) #removes the ASVs that are commonly enriched/depleted by the three strains

#now make vectors with ASVs significantly enriched/depleated by two strains
DA_WCS417_WCS315_v <- Reduce(intersect, list(DA_WCS417_vector,DA_WCS315_vector)) #322
DA_WCS417_WCS315_v <- setdiff(DA_WCS417_WCS315_v, DA_all_treat_vector) #61

DA_WCS417_CHA0_v <- Reduce(intersect, list(DA_WCS417_vector,DA_CHA0_vector)) #361
DA_WCS417_CHA0_v <- setdiff(DA_WCS417_CHA0_v, DA_all_treat_vector) #100

DA_WCS315_CHA0_v <- Reduce(intersect, list(DA_WCS315_vector,DA_CHA0_vector)) #283
DA_WCS315_CHA0_v <- setdiff(DA_WCS315_CHA0_v, DA_all_treat_vector) #22

DA_2strains_v <- c(DA_WCS417_WCS315_v, DA_WCS417_CHA0_v, DA_WCS315_CHA0_v) #183

#make a dataframe only with the ASVs in DA_2strains_v
df_2starins_f <- df_combined %>%
  filter(ASV %in% DA_2strains_v)

# Optional — remove prefixes like "p__", "f__", etc., for cleaner plot labels
clean_taxonomy <- function(x) {
  x <- trimws(x)  # Remove leading/trailing spaces
  gsub("^\\w__", "", x)  # Remove the prefix
}

# Apply the function with mutate
df_2starins_WW_f <- df_2starins_f %>%
  mutate(across(c(Phylum, Family, Genus), clean_taxonomy))

#DR

    ancombc_DR_l

WCS417_df <- ancombc_DR_l$df_sig_asv_l$WCS417 %>% 
  rownames_to_column(var = "ASV")

WCS315_df <- ancombc_DR_l$df_sig_asv_l$WCS315 %>% 
  rownames_to_column(var = "ASV")

CHA0_df <- ancombc_DR_l$df_sig_asv_l$CHA0 %>% 
  rownames_to_column(var = "ASV")

#join in a single dataframe
df_combined <- rbind(WCS417_df, WCS315_df, CHA0_df)


# What if I select only the ASVs that are significant for the 2 treatments

df_combined

#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- WCS417_df$ASV #695 ASVs
DA_WCS315_vector <- WCS315_df$ASV #630 ASVs
DA_CHA0_vector <- CHA0_df$ASV #386 ASVs

#I want a vector with only the ASVs  are differentially abundance under the three treatments when compared to control, so keep only ASVs that are present in the 3 vectors

DA_all_treat_vector <- Reduce(intersect, list(DA_WCS417_vector,DA_WCS315_vector,DA_CHA0_vector)) #261 ASVs

df_combined_f <- df_combined %>%
  filter(!ASV %in% DA_all_treat_vector) #removes the ASVs that are commonly enriched/depleted by the three strains

#now make vectors with ASVs significantly enriched/depleated by two strains
DA_WCS417_WCS315_v <- Reduce(intersect, list(DA_WCS417_vector,DA_WCS315_vector)) #285
DA_WCS417_WCS315_v <- setdiff(DA_WCS417_WCS315_v, DA_all_treat_vector) #65

DA_WCS417_CHA0_v <- Reduce(intersect, list(DA_WCS417_vector,DA_CHA0_vector)) #239
DA_WCS417_CHA0_v <- setdiff(DA_WCS417_CHA0_v, DA_all_treat_vector) #19

DA_WCS315_CHA0_v <- Reduce(intersect, list(DA_WCS315_vector,DA_CHA0_vector)) #246
DA_WCS315_CHA0_v <- setdiff(DA_WCS315_CHA0_v, DA_all_treat_vector) #26

DA_2strains_v <- c(DA_WCS417_WCS315_v, DA_WCS417_CHA0_v, DA_WCS315_CHA0_v) #183

#make a dataframe only with the ASVs in DA_2strains_v
df_2starins_f <- df_combined %>%
  filter(ASV %in% DA_2strains_v)

# Optional — remove prefixes like "p__", "f__", etc., for cleaner plot labels
clean_taxonomy <- function(x) {
  x <- trimws(x)  # Remove leading/trailing spaces
  gsub("^\\w__", "", x)  # Remove the prefix
}

# Apply the function with mutate
df_2starins_DR_f <- df_2starins_f %>%
  mutate(across(c(Phylum, Family, Genus), clean_taxonomy))

#join both WW and DR
df_2starins_WW_f <- df_2starins_WW_f %>% mutate(water_regime = "WW")
df_2starins_DR_f <- df_2starins_DR_f %>% mutate(water_regime = "DR")

#join the df

df_2starins_combined_f <- rbind(df_2starins_WW_f, df_2starins_DR_f)
#and if I order the phyllum based on more to less represented?

unique(df_2starins_combined_f$Phylum) #16 phylums

sum(df_2starins_combined_f$Phylum == "Planctomycetota", na.rm = TRUE) #52
sum(df_2starins_combined_f$Phylum == "Patescibacteria", na.rm = TRUE) #28
sum(df_2starins_combined_f$Phylum == "Proteobacteria", na.rm = TRUE) #290
sum(df_2starins_combined_f$Phylum == "Bacteroidota", na.rm = TRUE) #52
sum(df_2starins_combined_f$Phylum == "Actinobacteriota", na.rm = TRUE) #34
sum(df_2starins_combined_f$Phylum == "Acidobacteriota", na.rm = TRUE) #40
sum(df_2starins_combined_f$Phylum == "Verrucomicrobiota", na.rm = TRUE) #32
sum(df_2starins_combined_f$Phylum == "WPS-2", na.rm = TRUE) #2
sum(df_2starins_combined_f$Phylum == "Cyanobacteria", na.rm = TRUE) #4
sum(df_2starins_combined_f$Phylum == "Gemmatimonadota", na.rm = TRUE) #6
sum(df_2starins_combined_f$Phylum == "Armatimonadota", na.rm = TRUE) #2
sum(df_2starins_combined_f$Phylum == "Bdellovibrionota", na.rm = TRUE)#10
sum(df_2starins_combined_f$Phylum == "Spirochaetota", na.rm = TRUE) #4
sum(df_2starins_combined_f$Phylum == "Chloroflexi", na.rm = TRUE) #10
sum(df_2starins_combined_f$Phylum == "Myxococcota", na.rm = TRUE) #4
sum(df_2starins_combined_f$Phylum == "Firmicutes", na.rm = TRUE) #2

#organise factor for plots
df_2starins_combined_f$water_regime <- factor(df_2starins_combined_f$water_regime, c("WW", "DR"))

df_2starins_combined_f$Phylum <- factor(df_2starins_combined_f$Phylum, c("Proteobacteria", "Planctomycetota", "Bacteroidota", "Actinobacteriota",  "Acidobacteriota",  "Verrucomicrobiota", "Patescibacteria",  "Bdellovibrionota", "Chloroflexi", 
                                                       "Gemmatimonadota",   "Cyanobacteria", "Myxococcota", "Spirochaetota","Armatimonadota","Firmicutes",  "WPS-2"))

df_2starins_combined_f$Treatment <- factor(df_2starins_combined_f$Treatment, c("Control", "WCS417", "WCS315", "CHA0"))

#now plot


ggplot(df_2starins_combined_f,
       aes(x = Family, y = lfc, fill = Treatment)) +
   geom_point(aes(fill=Treatment), shape=21, color="black", stroke = 0.3,
             size=3, alpha=0.7)+ # individual sample values
  labs(y = "LFC", color = "Treatment") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  scale_fill_manual(values = c("#C8FA57", "#FA9F57", "#9771A5")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size=0.4)+
  facet_grid(water_regime~Phylum, scales = "free_x", space = "free_x")+
   theme_bw() + 
  theme(axis.text.y = element_text(colour = "black", size = 8, face = "bold"))+
  theme(axis.text.x = element_text(colour = "black", size = 8, angle = 90, hjust = 1, face = "plain"))

mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./results/plots/differential_abundance",
  scale = 1.8,
  width = 3500,
  height = 1250,
  units = "px",
  dpi = 300),
x=c("lfc_bulk_FINAL_2strains_March25.svg","lfc_bulk_FINAL_2strains_March25.png"))

```

## Heatmap2: Log-fold change per treatment

To get a joint heatmap (WCS417, WCS35 and CHA0 ) with log2fold differences between the treatments and control per ASV.
##WW
```{r}

#we will use the log2fold changes from ANCOM

ancombc_WW_l

  
#Keep only the information needed (ASVs and Lfc)
lfc_417 <- ancombc_WW_l$results$WCS417[,1, drop=FALSE]
lfc_315 <- ancombc_WW_l$results$WCS315[,1, drop=FALSE] 
lfc_CHA0 <- ancombc_WW_l$results$CHA0[,1, drop=FALSE] 


names(lfc_417) <- "lfc_WCS417"
names(lfc_315) <- "lfc_WCS315"
names(lfc_CHA0) <- "lfc_CHA0"

lfc_df <- cbind(lfc_417, lfc_315, lfc_CHA0) #merge

dim(lfc_df)

#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- rownames(ancombc_WW_l$df_sig_asv_l$WCS417)
DA_WCS315_vector <- rownames(ancombc_WW_l$df_sig_asv_l$WCS315)
DA_CHA0_vector <- rownames(ancombc_WW_l$df_sig_asv_l$CHA0)


DA_taxa <- c(DA_WCS417_vector, DA_WCS315_vector, DA_CHA0_vector) #join DA vectors
DA_taxa <-  unique(DA_taxa) #remove duplicate ASVs


DA_lfc_df <- lfc_df[DA_taxa, ] #matrix for pheatmap
length(DA_taxa) #1596
dim(DA_lfc_df) #1596

#metadada

DA_contrast_df <- DA_lfc_df 

DA_contrast_df <- DA_contrast_df %>% as.data.frame() %>% rownames_to_column(var = "ASV")

DA_contrast_df <- DA_contrast_df %>%
      mutate(DA_WCS417 = case_when(ASV %in% DA_WCS417_vector ~ "1",
                                TRUE ~ "0"))%>%
                mutate(DA_WCS315 = case_when(ASV %in% DA_WCS315_vector ~ "1",
                                TRUE ~ "0")) %>%
                mutate(DA_CHA0 = case_when(ASV %in% DA_CHA0_vector ~ "1",
                                TRUE ~ "0")) 


DA_contrast_df <- DA_contrast_df[, c(1, 5,6,7)]

DA_contrast_df <- DA_contrast_df %>%  column_to_rownames("ASV")


#### BULK SOIL -WW ####
matrix.lfc <- as.matrix(DA_lfc_df)

#make ps with DA asvs
ps_DA <- prune_taxa(DA_taxa, ps_WW_bulk_f)

#get species name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Genus")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Genus <- ifelse(is.na(df_row$Genus), df_row$asv_name, df_row$Genus)
df_row <- df_row %>%
           dplyr::select(Genus)


#change order DA_contrast_df 
DA_contrast_df_ordered <- DA_contrast_df[order(match(rownames(DA_contrast_df),
                                                     rownames(df_row))), , drop=FALSE]
#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]),
    Contrast = DA_contrast_df_ordered)
 
#change it to factor so it can be colored
annotation_row$Contrast.DA_WCS417 <- as.factor(annotation_row$Contrast.DA_WCS417)
annotation_row$Contrast.DA_WCS315 <- as.factor(annotation_row$Contrast.DA_WCS315)
annotation_row$Contrast.DA_CHA0 <- as.factor(annotation_row$Contrast.DA_CHA0)


#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.lfc)
rownames(annotation_row) #not in same order

#change order to matrix order 
matrix_row_ordered <- matrix.lfc[order(match(rownames(matrix.lfc),
                                                     rownames(annotation_row))), , drop=FALSE]
#check again
rownames(matrix_row_ordered)
rownames(annotation_row) #yes!

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.lfc))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row$Phylum %>% unique()
  length(Phyla) #24

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 24 colors for Phylum
display.brewer.pal(n = 24, name = 'Set3') #nice colors but we need more 

#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 24 #number of Phyla
color_phylum  <- colorRampPalette(brewer.pal(12, "Set3"))(colourCount) #vector with 15 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla$Phylum, color_phylum)
  
 
#put all colors together
annotation_colors = list(
     #Phylum = phylum_col,
     Contrast.DA_WCS417 = c("1" = "black", "0" = "white"),
     Contrast.DA_WCS315 = c("1" = "black", "0" = "white"),
     Contrast.DA_CHA0 = c("1" = "black", "0" = "white"))

#make colors for matrix
#Check first highest and ogest values in the matrix, in this case -5,+5
library(circlize)
function_color_matrix <- colorRamp2(c(-5, 0, 5), c("#018571", "white", "#A6611A")) #palette BrBG
function_color_matrix(seq(-5,5)) #gives you 16 colors with white in the middle 

# now plot the pheatmap
full_heatmap <- ComplexHeatmap::pheatmap(matrix_row_ordered,
                         width = ncol(matrix_row_ordered)*unit(5, "mm"), 
                         height = nrow(matrix_row_ordered)*unit(5, "mm"),
                         color= function_color_matrix,
                         name="LFC",
                         scale= "none", 
                         row_labels = species_ordered$Genus,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         main = "Treatment",
                         fontsize_row = 8,
                         fontsize_col = 8)

library(ggplotify)

gg_p <- as.ggplot(full_heatmap) 
gg_p


mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./results/plots/differential_abundance",
  scale = 1.8,
  width = 3000,
  height = 3000,
  units = "px",
  dpi = 300),
x=c("heatmap_WW_March25.svg","heatmap_WW_March25.png"))




```


##DR
```{r}

#we will use the log2fold changes from ANCOM

ancombc_DR_l

  
#Keep only the information needed (ASVs and Lfc)
lfc_417 <- ancombc_DR_l$results$WCS417[,1, drop=FALSE]
lfc_315 <- ancombc_DR_l$results$WCS315[,1, drop=FALSE] 
lfc_CHA0 <- ancombc_DR_l$results$CHA0[,1, drop=FALSE] 


names(lfc_417) <- "lfc_WCS417"
names(lfc_315) <- "lfc_WCS315"
names(lfc_CHA0) <- "lfc_CHA0"

lfc_df <- cbind(lfc_417, lfc_315, lfc_CHA0) #merge

dim(lfc_df)

#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- rownames(ancombc_DR_l$df_sig_asv_l$WCS417)
DA_WCS315_vector <- rownames(ancombc_DR_l$df_sig_asv_l$WCS315)
DA_CHA0_vector <- rownames(ancombc_DR_l$df_sig_asv_l$CHA0)


DA_taxa <- c(DA_WCS417_vector, DA_WCS315_vector, DA_CHA0_vector) #join DA vectors
DA_taxa <-  unique(DA_taxa) #remove duplicate ASVs


DA_lfc_df <- lfc_df[DA_taxa, ] #matrix for pheatmap
length(DA_taxa) #1161
dim(DA_lfc_df) #1161

#metadada

DA_contrast_df <- DA_lfc_df 

DA_contrast_df <- DA_contrast_df %>% as.data.frame() %>% rownames_to_column(var = "ASV")

DA_contrast_df <- DA_contrast_df %>%
      mutate(DA_WCS417 = case_when(ASV %in% DA_WCS417_vector ~ "1",
                                TRUE ~ "0"))%>%
                mutate(DA_WCS315 = case_when(ASV %in% DA_WCS315_vector ~ "1",
                                TRUE ~ "0")) %>%
                mutate(DA_CHA0 = case_when(ASV %in% DA_CHA0_vector ~ "1",
                                TRUE ~ "0")) 


DA_contrast_df <- DA_contrast_df[, c(1, 5,6,7)]

DA_contrast_df <- DA_contrast_df %>%  column_to_rownames("ASV")


#### BULK SOIL - DR ####
matrix.lfc <- as.matrix(DA_lfc_df)

#make ps with DA asvs
ps_DA <- prune_taxa(DA_taxa, ps_DR_bulk_f)

#get species name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Genus")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Genus <- ifelse(is.na(df_row$Genus), df_row$asv_name, df_row$Genus)
df_row <- df_row %>%
           dplyr::select(Genus)


#change order DA_contrast_df 
DA_contrast_df_ordered <- DA_contrast_df[order(match(rownames(DA_contrast_df),
                                                     rownames(df_row))), , drop=FALSE]
#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]),
    Contrast = DA_contrast_df_ordered)
 
#change it to factor so it can be colored
annotation_row$Contrast.DA_WCS417 <- as.factor(annotation_row$Contrast.DA_WCS417)
annotation_row$Contrast.DA_WCS315 <- as.factor(annotation_row$Contrast.DA_WCS315)
annotation_row$Contrast.DA_CHA0 <- as.factor(annotation_row$Contrast.DA_CHA0)


#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.lfc)
rownames(annotation_row) #not in same order

#change order to matrix order 
matrix_row_ordered <- matrix.lfc[order(match(rownames(matrix.lfc),
                                                     rownames(annotation_row))), , drop=FALSE]
#check again
rownames(matrix_row_ordered)
rownames(annotation_row) #yes!

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.lfc))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row$Phylum %>% unique()
  length(Phyla) #22

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 24 colors for Phylum
display.brewer.pal(n = 22, name = 'Set3') #nice colors but we need more 

#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 22 #number of Phyla
color_phylum  <- colorRampPalette(brewer.pal(12, "Set3"))(colourCount) #vector with 15 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla$Phylum, color_phylum)
  
 
#put all colors together
annotation_colors = list(
     #Phylum = phylum_col,
     Contrast.DA_WCS417 = c("1" = "black", "0" = "white"),
     Contrast.DA_WCS315 = c("1" = "black", "0" = "white"),
     Contrast.DA_CHA0 = c("1" = "black", "0" = "white"))

#make colors for matrix
#Check first highest and ogest values in the matrix, in this case -5,+5
library(circlize)
function_color_matrix <- colorRamp2(c(-5, 0, 5), c("#018571", "white", "#A6611A")) #palette BrBG
function_color_matrix(seq(-5,5)) #gives you 16 colors with white in the middle 

# now plot the pheatmap
full_heatmap_DR <- ComplexHeatmap::pheatmap(matrix_row_ordered,
                         width = ncol(matrix_row_ordered)*unit(5, "mm"), 
                         height = nrow(matrix_row_ordered)*unit(5, "mm"),
                         color= function_color_matrix,
                         name="LFC",
                         scale= "none", 
                         row_labels = species_ordered$Genus,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         main = "Treatment",
                         fontsize_row = 8,
                         fontsize_col = 8)

library(ggplotify)

gg_p <- as.ggplot(full_heatmap_DR) 
gg_p


mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./results/plots/differential_abundance",
  scale = 1.8,
  width = 3000,
  height = 3000,
  units = "px",
  dpi = 300),
x=c("heatmap_DR_March25.svg","heatmap_DR_March25.png"))




```

##data for venn diagrams
```{r}

## WW ##

#UP#
WW_WCS417_up <- ancombc_WW_l$df_sig_asv_l$WCS417 %>% filter(lfc>0) %>%     rownames_to_column("ASV_WCS417")

write.csv(WW_WCS417_up, "WW_WCS417_UP.csv", row.names = FALSE)

WW_WCS315_up <- ancombc_WW_l$df_sig_asv_l$WCS315 %>% filter(lfc>0) %>%     rownames_to_column("ASV_WCS315")

write.csv(WW_WCS315_up, "WW_WCS315_UP.csv", row.names = FALSE)

WW_CHA0_up <- ancombc_WW_l$df_sig_asv_l$CHA0 %>% filter(lfc>0)  %>%     rownames_to_column("ASV_CHA0")

write.csv(WW_CHA0_up, "WW_CHA0_UP.csv", row.names = FALSE)

#DOWN#

WW_WCS417_down <- ancombc_WW_l$df_sig_asv_l$WCS417 %>% filter(lfc<0) %>%     rownames_to_column("ASV_WCS417")

write.csv(WW_WCS417_down, "WW_WCS417_DOWN.csv", row.names = FALSE)

WW_WCS315_down <- ancombc_WW_l$df_sig_asv_l$WCS315 %>% filter(lfc<0) %>%     rownames_to_column("ASV_WCS315")

write.csv(WW_WCS315_down, "WW_WCS315_DOWN.csv", row.names = FALSE)

WW_CHA0_down <- ancombc_WW_l$df_sig_asv_l$CHA0 %>% filter(lfc<0) %>%     rownames_to_column("ASV_CHA0")

write.csv(WW_CHA0_down, "WW_CHA0_DOWN.csv", row.names = FALSE)



## DR ##

#UP#
DR_WCS417_up <- ancombc_DR_l$df_sig_asv_l$WCS417 %>% filter(lfc>0) %>%     rownames_to_column("ASV_WCS417")

write.csv(DR_WCS417_up, "DR_WCS417_UP.csv", row.names = FALSE)

DR_WCS315_up <- ancombc_DR_l$df_sig_asv_l$WCS315 %>% filter(lfc>0) %>%     rownames_to_column("ASV_WCS315")

write.csv(DR_WCS315_up, "DR_WCS315_UP.csv", row.names = FALSE)

DR_CHA0_up <- ancombc_DR_l$df_sig_asv_l$CHA0 %>% filter(lfc>0)  %>%     rownames_to_column("ASV_CHA0")

write.csv(DR_CHA0_up, "DR_CHA0_UP.csv", row.names = FALSE)

#DOWN#

DR_WCS417_down <- ancombc_DR_l$df_sig_asv_l$WCS417 %>% filter(lfc<0) %>%     rownames_to_column("ASV_WCS417")

write.csv(DR_WCS417_down, "DR_WCS417_DOWN.csv", row.names = FALSE)

DR_WCS315_down <- ancombc_DR_l$df_sig_asv_l$WCS315 %>% filter(lfc<0) %>%     rownames_to_column("ASV_WCS315")

write.csv(DR_WCS315_down, "DR_WCS315_DOWN.csv", row.names = FALSE)

DR_CHA0_down <- ancombc_DR_l$df_sig_asv_l$CHA0 %>% filter(lfc<0) %>%     rownames_to_column("ASV_CHA0")

write.csv(DR_CHA0_down, "DR_CHA0_DOWN.csv", row.names = FALSE)



```


##Significant for all

Now lets try to make a heatmap where we only show the ASVs that are significantly enriched/depleated for the 3 pseudomonas treatments

```{r}

#we will use the log2fold changes from ANCOM

ancombc_l

  
#Keep only the information needed (ASVs and Lfc)
lfc_417 <- ancombc_l$results$WCS417[,1, drop=FALSE]
lfc_315 <- ancombc_l$results$WCS315[,1, drop=FALSE] 
lfc_CHA0 <- ancombc_l$results$CHA0[,1, drop=FALSE] 


names(lfc_417) <- "lfc_WCS417"
names(lfc_315) <- "lfc_WCS315"
names(lfc_CHA0) <- "lfc_CHA0"

lfc_df <- cbind(lfc_417, lfc_315, lfc_CHA0) #merge

dim(lfc_df)

#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- rownames(ancombc_l$df_sig_asv_l$WCS417) #1000 ASVs
DA_WCS315_vector <- rownames(ancombc_l$df_sig_asv_l$WCS315) #1155 ASVs
DA_CHA0_vector <- rownames(ancombc_l$df_sig_asv_l$CHA0) #774 ASVs

#I want to keep only the ASVs  are differentially abundance under the three treatments when compared to control, so keep only ASVs that are present in the 3 vectors

DA_all_treat_vector <- Reduce(intersect, list(DA_WCS417_vector,DA_WCS315_vector,DA_CHA0_vector)) #354 ASVs


DA_lfc_df <- lfc_df[DA_all_treat_vector, ] #matrix for pheatmap
length(DA_all_treat_vector) #354
dim(DA_lfc_df) #354


#### Rhizosphere ####
matrix.lfc <- as.matrix(DA_lfc_df)

#make ps with DA asvs
ps_DA <- prune_taxa(DA_all_treat_vector, ps_WW_rhizo_f)

#get species name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Genus")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Genus <- ifelse(is.na(df_row$Genus), df_row$asv_name, df_row$Genus)
df_row <- df_row %>%
           dplyr::select(Genus)


#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]))



#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.lfc)
rownames(annotation_row) #not in same order

#change order to matrix order 
matrix_row_ordered <- matrix.lfc[order(match(rownames(matrix.lfc),
                                                     rownames(annotation_row))), , drop=FALSE]
#check again
rownames(matrix_row_ordered)
rownames(annotation_row) #yes!

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.lfc))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row$Phylum %>% unique()
  length(Phyla) #17

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 30 colors for Phylum
display.brewer.pal(n = 17, name = 'Set3') #nice colors but we need more 

#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 17 #number of Phyla
color_phylum  <- colorRampPalette(brewer.pal(12, "Set3"))(colourCount) #vector with 15 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla$Phylum, color_phylum)
  
 

#make colors for matrix
#Check first highest and ogest values in the matrix, in this case -4,+4
library(circlize)
function_color_matrix <- colorRamp2(c(-4, 0, 4), c("#018571", "white", "#A6611A")) #palette BrBG
function_color_matrix(seq(-4,4)) #gives you 8 colors with white in the middle 

# now plot the pheatmap
heatmap_ancombc_sig_all_treatments <- ComplexHeatmap::pheatmap(matrix_row_ordered,
                         color= function_color_matrix,
                         name="LFC",
                         scale= "none", 
                         row_labels = species_ordered$Genus,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         #annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         main = "Treatment",
                         fontsize_row = 8,
                         fontsize_col = 8)











#### What if I use a cut-off of lfc> + - 2 ####

#we will use the log2fold changes from ANCOM

ancombc_l

  
#Keep only the information needed (ASVs and Lfc)
lfc_417 <- ancombc_l$results$WCS417[,1, drop=FALSE]
lfc_315 <- ancombc_l$results$WCS315[,1, drop=FALSE] 
lfc_CHA0 <- ancombc_l$results$CHA0[,1, drop=FALSE] 


names(lfc_417) <- "lfc_WCS417"
names(lfc_315) <- "lfc_WCS315"
names(lfc_CHA0) <- "lfc_CHA0"

lfc_df <- cbind(lfc_417, lfc_315, lfc_CHA0) #merge

#keep only ASVs that are >= +-2 for one of the treatments 

enriched <- lfc_df[apply(lfc_df > 2,1,any),] #311
depleated <- lfc_df[apply(lfc_df < -2,1,any),] #229

lfc_2_df <- rbind(enriched, depleated) #540

dim(lfc_2_df) #540

#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- rownames(ancombc_l$df_sig_asv_l$WCS417) #1000 ASVs
DA_WCS315_vector <- rownames(ancombc_l$df_sig_asv_l$WCS315) #1155 ASVs
DA_CHA0_vector <- rownames(ancombc_l$df_sig_asv_l$CHA0) #774 ASVs

#I want to keep only the ASVs  are differentially abundance under the three treatments when compared to control, so keep only ASVs that are present in the 3 vectors

DA_all_treat_vector <- Reduce(intersect, list(DA_WCS417_vector,DA_WCS315_vector,DA_CHA0_vector)) #354 ASVs

#Keep only the DA that have lfc > 2 -2 

lfc_2_vector <- rownames(lfc_2_df) #vector of ASV with lfc >= +2 <= -2

DA_lfc_2_vector <- intersect(DA_all_treat_vector, lfc_2_vector)

DA_lfc_2_df <- lfc_2_df[DA_lfc_2_vector, ] #matrix for pheatmap
length(DA_lfc_2_vector) #85
dim(DA_lfc_2_df) #85


#### Rhizosphere ####
matrix.lfc <- as.matrix(DA_lfc_2_df)

#make ps with DA asvs and lfc cut off
ps_DA <- prune_taxa(DA_lfc_2_vector, ps_WW_rhizo_f)

#get species name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Genus")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Genus <- ifelse(is.na(df_row$Genus), df_row$asv_name, df_row$Genus)
df_row <- df_row %>%
           dplyr::select(Genus)


#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]))



#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.lfc)
rownames(annotation_row) #same order

#change order to matrix order 
matrix_row_ordered <- matrix.lfc[order(match(rownames(matrix.lfc),
                                                     rownames(annotation_row))), , drop=FALSE]
#check again
rownames(matrix_row_ordered)
rownames(annotation_row) #yes!


#change order to matrix order 
matrix_row_ordered <- matrix.lfc[order(match(rownames(matrix.lfc),
                                                     rownames(annotation_row))), , drop=FALSE]

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.lfc))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row$Phylum %>% unique()
  length(Phyla) #14

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 14 colors for Phylum
display.brewer.pal(n = 14, name = 'Set3') #nice colors but we need more 


#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 14 #number of Phyla
color_phylum  <- colorRampPalette(brewer.pal(12, "Set3"))(colourCount) #vector with 14 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla, color_phylum)
  
 

#make colors for matrix
#Check first highest and ogest values in the matrix, in this case -4,+4
library(circlize)
function_color_matrix <- colorRamp2(c(-4, 0, 4), c("#018571", "white", "#A6611A")) #palette BrBG
function_color_matrix(seq(-4,4)) #gives you 8 colors with white in the middle 

# now plot the pheatmap
heatmap_ancombc_sig_all_lfc2_treatments <- ComplexHeatmap::pheatmap(matrix_row_ordered,
                         color= function_color_matrix,
                         name="LFC",
                         scale= "none", 
                         row_labels = species_ordered$Genus,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                        #annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         main = "Treatment",
                         fontsize_row = 15,
                         fontsize_col = 15)





```

##DA in WCS417 and CHA0, but not in WCS315

```{r}

#we will use the log2fold changes from ANCOM

ancombc_l

  
#Keep only the information needed (ASVs and Lfc)
lfc_417 <- ancombc_l$results$WCS417[,1, drop=FALSE]
lfc_315 <- ancombc_l$results$WCS315[,1, drop=FALSE] 
lfc_CHA0 <- ancombc_l$results$CHA0[,1, drop=FALSE] 


names(lfc_417) <- "lfc_WCS417"
names(lfc_315) <- "lfc_WCS315"
names(lfc_CHA0) <- "lfc_CHA0"

lfc_df <- cbind(lfc_417, lfc_315, lfc_CHA0) #merge

lfc_417_CHA0_df <- cbind(lfc_417, lfc_CHA0) #merge

dim(lfc_df) #6577
dim(lfc_417_CHA0_df) #6577

#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- rownames(ancombc_l$df_sig_asv_l$WCS417) #1000 ASVs
DA_WCS315_vector <- rownames(ancombc_l$df_sig_asv_l$WCS315) #1155 ASVs
DA_CHA0_vector <- rownames(ancombc_l$df_sig_asv_l$CHA0) #774 ASVs

#I want to keep only the ASVs  are differentially abundance under fo WCS417 and CHA0 to control, but not for WCS315

DA_417_CHA0_vector <- intersect(DA_WCS417_vector,DA_CHA0_vector) #447 ASVs
remove_ASV_vector <- intersect(DA_417_CHA0_vector, DA_WCS315_vector) #354 --> ASVs I wont to remove
DA_417_CHA0_vector <- outersect(DA_417_CHA0_vector, remove_ASV_vector) #93



DA_lfc_417_CHA0_df <- lfc_417_CHA0_df[DA_417_CHA0_vector, ] #matrix for pheatmap
length(DA_417_CHA0_vector) #93
dim(DA_lfc_417_CHA0_df) #93


#### Rhizosphere ####
matrix.lfc <- as.matrix(DA_lfc_417_CHA0_df)

#make ps with DA asvs
ps_DA <- prune_taxa(DA_417_CHA0_vector, ps_WW_rhizo_f)

#get species name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Genus")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Genus <- ifelse(is.na(df_row$Genus), df_row$asv_name, df_row$Genus)
df_row <- df_row %>%
           dplyr::select(Genus)


#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]))



#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.lfc)
rownames(annotation_row) #not in same order

#change order to matrix order 
matrix_row_ordered <- matrix.lfc[order(match(rownames(matrix.lfc),
                                                     rownames(annotation_row))), , drop=FALSE]
#check again
rownames(matrix_row_ordered)
rownames(annotation_row) #yes!

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.lfc))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row$Phylum %>% unique()
  length(Phyla) #12

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 30 colors for Phylum
display.brewer.pal(n = 12, name = 'Set3') #nice colors but we need more 

#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 12 #number of Phyla
color_phylum  <- brewer.pal(12, "Set3") #vector with 15 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla, color_phylum)


#define annotation colors
annotation_colors <- phylum_col
 

#make colors for matrix
#Check first highest and ogest values in the matrix, in this case -4,+4
library(circlize)
function_color_matrix <- colorRamp2(c(-4, 0, 4), c("#018571", "white", "#A6611A")) #palette BrBG
function_color_matrix(seq(-4,4)) #gives you 8 colors with white in the middle 

# now plot the pheatmap
heatmap_uniqueDA417CHA0 <- ComplexHeatmap::pheatmap(matrix_row_ordered,
                         color= function_color_matrix,
                         name="LFC",
                         scale= "none", 
                         row_labels = species_ordered$Genus,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         main = "Treatment",
                         fontsize_row = 15,
                         fontsize_col = 15)







##### if we use a cutoff of  lfc 2 ####

matrix.lfc <- as.matrix(DA_lfc_417_CHA0_df)

enriched <- matrix.lfc[apply(matrix.lfc > 2,1,any),] #10
depleated <- matrix.lfc[apply(matrix.lfc < -2,1,any),] #22

matrix.lfc2 <- rbind(enriched, depleated) #32

DA_417_CHA0__lfc2_v <- rownames(matrix.lfc2)

#make ps with DA asvs
ps_DA <- prune_taxa(DA_417_CHA0__lfc2_v, ps_WW_rhizo_f)

#get species name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Genus")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Genus <- ifelse(is.na(df_row$Genus), df_row$asv_name, df_row$Genus)
df_row <- df_row %>%
           dplyr::select(Genus)


#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]))



#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.lfc2)
rownames(annotation_row) #not in same order

#change order to matrix order 
matrix_row_ordered <- matrix.lfc2[order(match(rownames(matrix.lfc2),
                                                     rownames(annotation_row))), , drop=FALSE]
#check again
rownames(matrix_row_ordered)
rownames(annotation_row) #yes!

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.lfc2))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row$Phylum %>% unique()
  length(Phyla) #7

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 30 colors for Phylum
display.brewer.pal(n = 7, name = 'Set3') #nice colors but we need more 

#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 7 #number of Phyla
color_phylum  <- brewer.pal(7, "Set3") #vector with 15 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla, color_phylum)


#define annotation colors
annotation_colors <- phylum_col
 

#make colors for matrix
#Check first highest and ogest values in the matrix, in this case -4,+4
library(circlize)
function_color_matrix <- colorRamp2(c(-4, 0, 4), c("#018571", "white", "#A6611A")) #palette BrBG
function_color_matrix(seq(-4,4)) #gives you 8 colors with white in the middle 

# now plot the pheatmap
 heatmap_uniqueDA417CHA0_lfc2 <- ComplexHeatmap::pheatmap(matrix_row_ordered,
                         color= function_color_matrix,
                         name="LFC",
                         scale= "none", 
                         row_labels = species_ordered$Genus,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         #annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         main = "Treatment",
                         fontsize_row = 15,
                         fontsize_col = 15)



```


```{r}
#### Rhizosphere ####
matrix.log2 <- as.matrix(DA_log2fold_df)

#make ps with DA asvs
ps_DA <- prune_taxa(DA_taxa, ps_WW_rhizo_f)

#get Family name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Family")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Family <- ifelse(is.na(df_row$Family), df_row$asv_name, df_row$Family)
df_row <- df_row %>%
           dplyr::select(Family)


#change order DA_contrast_df
DA_contrast_df_ordered <- DA_contrast_df[order(match(rownames(DA_contrast_df),
                                                     rownames(annotation_row))), , drop=FALSE]
#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]),
    Contrast = DA_contrast_df_ordered)
 
#change it to factor so it can be colored
annotation_row$Contrast.DA_WCS417 <- as.factor(annotation_row$Contrast.DA_WCS417)
annotation_row$Contrast.DA_WCS315 <- as.factor(annotation_row$Contrast.DA_WCS315)
annotation_row$Contrast.DA_CHA0 <- as.factor(annotation_row$Contrast.DA_CHA0)


#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.log2)
rownames(annotation_row) #not in same order

#change order to matrix order 
matrix_row_ordered <- matrix.log2[order(match(rownames(matrix.log2),
                                                     rownames(annotation_row))), , drop=FALSE]
#check again
rownames(matrix_row_ordered)
rownames(annotation_row) #yes!

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.log2))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row %>% #17
         count(Phylum)

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 17 colors for Phylum
display.brewer.pal(n = 17, name = 'Set3') #nice colors but we need more 

#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 17 #number of Phyla
color_phylum  <- colorRampPalette(brewer.pal(12, "Set3"))(colourCount) #vector with 15 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla$Phylum, color_phylum)
  
 
#put all colors together
annotation_colors = list(
     #Phylum = phylum_col,
     Contrast.DA_WCS417 = c("1" = "black", "0" = "white"),
     Contrast.DA_WCS315 = c("1" = "black", "0" = "white"),
     Contrast.DA_CHA0 = c("1" = "black", "0" = "white"))

#make colors for matrix
library(circlize)
function_color_matrix <- colorRamp2(c(-7, 0, 7), c("#018571", "white", "#A6611A")) #palette BrBG
function_color_matrix(seq(-7,7)) #gives you 16 colors with white in the middle 

# now plot the pheatmap
ComplexHeatmap::pheatmap(matrix_row_ordered,
                         color= function_color_matrix,
                         name="L2FC",
                         scale= "none", 
                         row_labels = species_ordered$Family,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         #cluster_rows=TRUE,
                         main = "Treatment",
                         fontsize_row = 8,
                         fontsize_col = 8)



```

## Venn diagrams

```{r}
ancombc_l



#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- rownames(ancombc_l$df_sig_asv_l$WCS417) #1000
DA_WCS315_vector <- rownames(ancombc_l$df_sig_asv_l$WCS315) #1155
DA_CHA0_vector <- rownames(ancombc_l$df_sig_asv_l$CHA0) #774

#With package  nVennR
library("nVennR")
library("rsvg")
library("grImport2")


myplot <- plotVenn(list(PsWCS417 = DA_WCS417_vector, 
           PsWCS315 = DA_WCS315_vector,
           PsCHA0 = DA_CHA0_vector),
          opacity = 0.4, borderWidth = 5,labelRegions = F, fontScale=3, 
          setColors = c("#A6CEE3", "#B2DF8A", "#CAB2D6"))


#Only when significant in all three and lfc>2



```



### What if we use only the robust ASVs

```{r}
#we will use the log2fold changes from normal deseq

deseq_norm_l

#Now, correct by shrienking with Log2 fold change 
  lfc_dds_417 <- lfcShrink(deseq_norm_l$result_test, coef=2, type="normal") 
  lfc_dds_315 <- lfcShrink(deseq_norm_l$result_test, coef=3, type="normal") 
  lfc_dds_CHA0 <- lfcShrink(deseq_norm_l$result_test, coef=4, type="normal") 
  
#Keep only the information needed (ASVs and log2fold)
log2fold_417 <- lfc_dds_417[2] 
log2fold_315 <- lfc_dds_315[2]
log2fold_CHA0 <- lfc_dds_CHA0[2]

names(log2fold_417) <- "log2fold_WCS417"
names(log2fold_315) <- "log2fold_WCS315"
names(log2fold_CHA0) <- "log2fold_CHA0"

log2fold_df <- cbind(log2fold_417, log2fold_315, log2fold_CHA0) #merge

dim(log2fold_df)

#Keep only log2fold of DA taxa (from all the treatments pull)


DA_WCS417_vector <- rownames(ps_robust_l$WCS417@tax_table)
DA_WCS315_vector <- rownames(ps_robust_l$wcs315@tax_table)
DA_CHA0_vector <- rownames(ps_robust_l$CHA0@tax_table)

DA_taxa <- c(DA_WCS417_vector, DA_WCS315_vector, DA_CHA0_vector) #join DA vectors
DA_taxa <-  unique(DA_taxa) #remove duplicate ASVs


DA_log2fold_df <- log2fold_df[DA_taxa, ] #matrix for pheatmap
dim(DA_log2fold_df)

#metadada

DA_contrast_df <- DA_log2fold_df 

DA_contrast_df <- DA_contrast_df %>% as.data.frame() %>% rownames_to_column(var = "ASV")

DA_contrast_df <- DA_contrast_df %>%
      mutate(DA_WCS417 = case_when(ASV %in% DA_WCS417_vector ~ "1",
                                TRUE ~ "0"))%>%
                mutate(DA_WCS315 = case_when(ASV %in% DA_WCS315_vector ~ "1",
                                TRUE ~ "0")) %>%
                mutate(DA_CHA0 = case_when(ASV %in% DA_CHA0_vector ~ "1",
                                TRUE ~ "0")) 


DA_contrast_df <- DA_contrast_df[, c(1, 5,6,7)]

DA_contrast_df <- DA_contrast_df %>%  column_to_rownames("ASV")




#### Rhizosphere ####
matrix.log2 <- as.matrix(DA_log2fold_df)

#make ps with DA asvs
ps_DA <- prune_taxa(DA_taxa, ps_WW_rhizo_f)

#get species name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Genus")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Genus <- ifelse(is.na(df_row$Genus), df_row$asv_name, df_row$Genus)
df_row <- df_row %>%
           dplyr::select(Genus)


#change order DA_contrast_df
DA_contrast_df_ordered <- DA_contrast_df[order(match(rownames(DA_contrast_df),
                                                     rownames(annotation_row))), , drop=FALSE]
#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]),
    Contrast = DA_contrast_df_ordered)
 
#change it to factor so it can be colored
annotation_row$Contrast.DA_WCS417 <- as.factor(annotation_row$Contrast.DA_WCS417)
annotation_row$Contrast.DA_WCS315 <- as.factor(annotation_row$Contrast.DA_WCS315)
annotation_row$Contrast.DA_CHA0 <- as.factor(annotation_row$Contrast.DA_CHA0)


#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.log2)
rownames(annotation_row) #not in same order

#change order to matrix order 
matrix_row_ordered <- matrix.log2[order(match(rownames(matrix.log2),
                                                     rownames(annotation_row))), , drop=FALSE]
#check again
rownames(matrix_row_ordered)
rownames(annotation_row) #yes!

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.log2))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row %>% #17
         count(Phylum)

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 17 colors for Phylum
display.brewer.pal(n = 17, name = 'Set3') #nice colors but we need more 

#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 17 #number of Phyla
color_phylum  <- colorRampPalette(brewer.pal(12, "Set3"))(colourCount) #vector with 15 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla$Phylum, color_phylum)
  
 
#put all colors together
annotation_colors = list(
     #Phylum = phylum_col,
     Contrast.DA_WCS417 = c("1" = "black", "0" = "white"),
     Contrast.DA_WCS315 = c("1" = "black", "0" = "white"),
     Contrast.DA_CHA0 = c("1" = "black", "0" = "white"))

#make colors for matrix
library(circlize)
function_color_matrix <- colorRamp2(c(-7, 0, 7), c("#018571", "white", "#A6611A")) #palette BrBG
function_color_matrix(seq(-7,7)) #gives you 16 colors with white in the middle 

# now plot the pheatmap
ComplexHeatmap::pheatmap(matrix_row_ordered,
                         color= function_color_matrix,
                         name="L2FC",
                         scale= "none", 
                         row_labels = species_ordered$Genus,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         #cluster_rows=TRUE,
                         main = "Treatment",
                         fontsize_row = 8,
                         fontsize_col = 20)

```

### Table with most importat ASVs

```{r}
matrix_row_ordered

#let's order the ASVs according to the ones with highest log2fold for WCS417
#change it to factor so it can be colored

DA_log2fold_df <- as.data.frame(DA_log2fold_df)
DA_log2fold_df <- DA_log2fold_df %>% arrange(desc(log2fold_WCS417))
matrix_row_ordered$log2fold_WCS417 <- as.numeric(matrix_row_ordered$log2fold_WCS417)
matrix_row_ordered <- matrix_row_ordered %>% arrange(desc(log2fold_WCS417))

str(matrix_row_ordered)





#Take our dataframe that tells as which ASV is signif DA under each treatment 
DA_contrast_df

#Add a new column that tell as under how many treatments the ASV is DA
DA_contrast_df$DA_WCS417 <- as.numeric(DA_contrast_df$DA_WCS417)
DA_contrast_df$DA_WCS315 <- as.numeric(DA_contrast_df$DA_WCS315)
DA_contrast_df$DA_CHA0 <- as.numeric(DA_contrast_df$DA_CHA0)

DA_contrast_df$n <-  rowSums(DA_contrast_df)

#Keep ASVs that are significant with 3 treatments
DA_contrast_all_sig_df <- filter(DA_contrast_df, n=="3")
dim(DA_contrast_all_sig_df) #17 ASVs are significant with 3 treatments

#Make a vector with the names of the 17 ASVs
all_sig_ASVs_v <- rownames(DA_contrast_all_sig_df)

#Check the log2fold of the 17 ASV
all_sig_DA_log2fold_df <- DA_log2fold_df[all_sig_ASVs_v,]

###Prepare for pheatmap###

matrix.log2 <- as.matrix(all_sig_DA_log2fold_df)

#make ps with DA asvs
ps_DA <- prune_taxa(all_sig_ASVs_v, ps_WW_rhizo_f)

#get species name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Genus")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Genus <- ifelse(is.na(df_row$Genus), df_row$asv_name, df_row$Genus)
df_row <- df_row %>%
           dplyr::select(Genus)


 #change order DA_contrast_df
DA_contrast_df_ordered <- DA_contrast_df[order(match(rownames(DA_contrast_df),
                                                     rownames(annotation_row))), , drop=FALSE]
#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]))
 
#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.log2)
rownames(annotation_row) #not in same order

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.log2))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row %>% #5
         count(Phylum)

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 17 colors for Phylum
display.brewer.pal(n = 5, name = 'Set3') #nice colors but we need more 

#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 5 #number of Phyla
color_phylum  <- colorRampPalette(brewer.pal(12, "Set3"))(colourCount) #vector with 15 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla$Phylum, color_phylum)
  
  # now plot the pheatmap
ComplexHeatmap::pheatmap(matrix.log2,
                         color= function_color_matrix,
                         name="L2FC",
                         scale= "none", 
                         row_labels = species_ordered$Genus,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         #annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         #cluster_rows=TRUE,
                         #main = "Treatment",
                         fontsize_row = 8,
                         fontsize_col = 10)


```





## Heatmap2: Log-fold change per treatment

To get a joint heatmap (WCS417, WCS35 and CHA0 ) with log2fold differences between the treatments and control per ASV.

```{r}

#we will use the log2fold changes from ANCOM

ancombc_l

  
#Keep only the information needed (ASVs and Lfc)
lfc_417 <- ancombc_l$results$WCS417[,1, drop=FALSE]
lfc_315 <- ancombc_l$results$WCS315[,1, drop=FALSE] 
lfc_CHA0 <- ancombc_l$results$CHA0[,1, drop=FALSE] 


names(lfc_417) <- "lfc_WCS417"
names(lfc_315) <- "lfc_WCS315"
names(lfc_CHA0) <- "lfc_CHA0"

lfc_df <- cbind(lfc_417, lfc_315, lfc_CHA0) #merge

dim(lfc_df)

#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- rownames(ancombc_l$df_sig_asv_l$WCS417)
DA_WCS315_vector <- rownames(ancombc_l$df_sig_asv_l$WCS315)
DA_CHA0_vector <- rownames(ancombc_l$df_sig_asv_l$CHA0)


DA_taxa <- c(DA_WCS417_vector, DA_WCS315_vector, DA_CHA0_vector) #join DA vectors
DA_taxa <-  unique(DA_taxa) #remove duplicate ASVs


DA_lfc_df <- lfc_df[DA_taxa, ] #matrix for pheatmap
length(DA_taxa) #1973
dim(DA_lfc_df) #1973

#metadada

DA_contrast_df <- DA_lfc_df 

DA_contrast_df <- DA_contrast_df %>% as.data.frame() %>% rownames_to_column(var = "ASV")

DA_contrast_df <- DA_contrast_df %>%
      mutate(DA_WCS417 = case_when(ASV %in% DA_WCS417_vector ~ "1",
                                TRUE ~ "0"))%>%
                mutate(DA_WCS315 = case_when(ASV %in% DA_WCS315_vector ~ "1",
                                TRUE ~ "0")) %>%
                mutate(DA_CHA0 = case_when(ASV %in% DA_CHA0_vector ~ "1",
                                TRUE ~ "0")) 


DA_contrast_df <- DA_contrast_df[, c(1, 5,6,7)]

DA_contrast_df <- DA_contrast_df %>%  column_to_rownames("ASV")


#### Rhizosphere ####
matrix.lfc <- as.matrix(DA_lfc_df)

#make ps with DA asvs
ps_DA <- prune_taxa(DA_taxa, ps_WW_rhizo_f)

#get species name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Genus")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Genus <- ifelse(is.na(df_row$Genus), df_row$asv_name, df_row$Genus)
df_row <- df_row %>%
           dplyr::select(Genus)


#change order DA_contrast_df 
DA_contrast_df_ordered <- DA_contrast_df[order(match(rownames(DA_contrast_df),
                                                     rownames(df_row))), , drop=FALSE]
#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]),
    Contrast = DA_contrast_df_ordered)
 
#change it to factor so it can be colored
annotation_row$Contrast.DA_WCS417 <- as.factor(annotation_row$Contrast.DA_WCS417)
annotation_row$Contrast.DA_WCS315 <- as.factor(annotation_row$Contrast.DA_WCS315)
annotation_row$Contrast.DA_CHA0 <- as.factor(annotation_row$Contrast.DA_CHA0)


#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.lfc)
rownames(annotation_row) #not in same order

#change order to matrix order 
matrix_row_ordered <- matrix.lfc[order(match(rownames(matrix.lfc),
                                                     rownames(annotation_row))), , drop=FALSE]
#check again
rownames(matrix_row_ordered)
rownames(annotation_row) #yes!

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.lfc))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row$Phylum %>% unique()
  length(Phyla) #30

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 30 colors for Phylum
display.brewer.pal(n = 30, name = 'Set3') #nice colors but we need more 

#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 17 #number of Phyla
color_phylum  <- colorRampPalette(brewer.pal(12, "Set3"))(colourCount) #vector with 15 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla$Phylum, color_phylum)
  
 
#put all colors together
annotation_colors = list(
     #Phylum = phylum_col,
     Contrast.DA_WCS417 = c("1" = "black", "0" = "white"),
     Contrast.DA_WCS315 = c("1" = "black", "0" = "white"),
     Contrast.DA_CHA0 = c("1" = "black", "0" = "white"))

#make colors for matrix
#Check first highest and ogest values in the matrix, in this case -5,+5
library(circlize)
function_color_matrix <- colorRamp2(c(-5, 0, 5), c("#018571", "white", "#A6611A")) #palette BrBG
function_color_matrix(seq(-5,5)) #gives you 16 colors with white in the middle 

# now plot the pheatmap
full_heatmap <- ComplexHeatmap::pheatmap(matrix_row_ordered,
                         width = ncol(matrix_row_ordered)*unit(5, "mm"), 
                         height = nrow(matrix_row_ordered)*unit(5, "mm"),
                         color= function_color_matrix,
                         name="LFC",
                         scale= "none", 
                         row_labels = species_ordered$Genus,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         main = "Treatment",
                         fontsize_row = 8,
                         fontsize_col = 8)








```

##Significant for all

Now lets try to make a heatmap where we only show the ASVs that are significantly enriched/depleated for the 3 pseudomonas treatments

```{r}

#we will use the log2fold changes from ANCOM

ancombc_l

  
#Keep only the information needed (ASVs and Lfc)
lfc_417 <- ancombc_l$results$WCS417[,1, drop=FALSE]
lfc_315 <- ancombc_l$results$WCS315[,1, drop=FALSE] 
lfc_CHA0 <- ancombc_l$results$CHA0[,1, drop=FALSE] 


names(lfc_417) <- "lfc_WCS417"
names(lfc_315) <- "lfc_WCS315"
names(lfc_CHA0) <- "lfc_CHA0"

lfc_df <- cbind(lfc_417, lfc_315, lfc_CHA0) #merge

dim(lfc_df)

#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- rownames(ancombc_l$df_sig_asv_l$WCS417) #1000 ASVs
DA_WCS315_vector <- rownames(ancombc_l$df_sig_asv_l$WCS315) #1155 ASVs
DA_CHA0_vector <- rownames(ancombc_l$df_sig_asv_l$CHA0) #774 ASVs

#I want to keep only the ASVs  are differentially abundance under the three treatments when compared to control, so keep only ASVs that are present in the 3 vectors

DA_all_treat_vector <- Reduce(intersect, list(DA_WCS417_vector,DA_WCS315_vector,DA_CHA0_vector)) #354 ASVs


DA_lfc_df <- lfc_df[DA_all_treat_vector, ] #matrix for pheatmap
length(DA_all_treat_vector) #354
dim(DA_lfc_df) #354


#### Rhizosphere ####
matrix.lfc <- as.matrix(DA_lfc_df)

#make ps with DA asvs
ps_DA <- prune_taxa(DA_all_treat_vector, ps_WW_rhizo_f)

#get species name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Genus")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Genus <- ifelse(is.na(df_row$Genus), df_row$asv_name, df_row$Genus)
df_row <- df_row %>%
           dplyr::select(Genus)


#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]))



#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.lfc)
rownames(annotation_row) #not in same order

#change order to matrix order 
matrix_row_ordered <- matrix.lfc[order(match(rownames(matrix.lfc),
                                                     rownames(annotation_row))), , drop=FALSE]
#check again
rownames(matrix_row_ordered)
rownames(annotation_row) #yes!

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.lfc))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row$Phylum %>% unique()
  length(Phyla) #17

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 30 colors for Phylum
display.brewer.pal(n = 17, name = 'Set3') #nice colors but we need more 

#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 17 #number of Phyla
color_phylum  <- colorRampPalette(brewer.pal(12, "Set3"))(colourCount) #vector with 15 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla$Phylum, color_phylum)
  
 

#make colors for matrix
#Check first highest and ogest values in the matrix, in this case -4,+4
library(circlize)
function_color_matrix <- colorRamp2(c(-4, 0, 4), c("#018571", "white", "#A6611A")) #palette BrBG
function_color_matrix(seq(-4,4)) #gives you 8 colors with white in the middle 

# now plot the pheatmap
heatmap_ancombc_sig_all_treatments <- ComplexHeatmap::pheatmap(matrix_row_ordered,
                         color= function_color_matrix,
                         name="LFC",
                         scale= "none", 
                         row_labels = species_ordered$Genus,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         #annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         main = "Treatment",
                         fontsize_row = 8,
                         fontsize_col = 8)











#### What if I use a cut-off of lfc> + - 2 ####

#we will use the log2fold changes from ANCOM

ancombc_l

  
#Keep only the information needed (ASVs and Lfc)
lfc_417 <- ancombc_l$results$WCS417[,1, drop=FALSE]
lfc_315 <- ancombc_l$results$WCS315[,1, drop=FALSE] 
lfc_CHA0 <- ancombc_l$results$CHA0[,1, drop=FALSE] 


names(lfc_417) <- "lfc_WCS417"
names(lfc_315) <- "lfc_WCS315"
names(lfc_CHA0) <- "lfc_CHA0"

lfc_df <- cbind(lfc_417, lfc_315, lfc_CHA0) #merge

#keep only ASVs that are >= +-2 for one of the treatments 

enriched <- lfc_df[apply(lfc_df > 2,1,any),] #311
depleated <- lfc_df[apply(lfc_df < -2,1,any),] #229

lfc_2_df <- rbind(enriched, depleated) #540

dim(lfc_2_df) #540

#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- rownames(ancombc_l$df_sig_asv_l$WCS417) #1000 ASVs
DA_WCS315_vector <- rownames(ancombc_l$df_sig_asv_l$WCS315) #1155 ASVs
DA_CHA0_vector <- rownames(ancombc_l$df_sig_asv_l$CHA0) #774 ASVs

#I want to keep only the ASVs  are differentially abundance under the three treatments when compared to control, so keep only ASVs that are present in the 3 vectors

DA_all_treat_vector <- Reduce(intersect, list(DA_WCS417_vector,DA_WCS315_vector,DA_CHA0_vector)) #354 ASVs

#Keep only the DA that have lfc > 2 -2 

lfc_2_vector <- rownames(lfc_2_df) #vector of ASV with lfc >= +2 <= -2

DA_lfc_2_vector <- intersect(DA_all_treat_vector, lfc_2_vector)

DA_lfc_2_df <- lfc_2_df[DA_lfc_2_vector, ] #matrix for pheatmap
length(DA_lfc_2_vector) #85
dim(DA_lfc_2_df) #85


#### Rhizosphere ####
matrix.lfc <- as.matrix(DA_lfc_2_df)

#make ps with DA asvs and lfc cut off
ps_DA <- prune_taxa(DA_lfc_2_vector, ps_WW_rhizo_f)

#get species name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Genus")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Genus <- ifelse(is.na(df_row$Genus), df_row$asv_name, df_row$Genus)
df_row <- df_row %>%
           dplyr::select(Genus)


#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]))



#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.lfc)
rownames(annotation_row) #same order

#change order to matrix order 
matrix_row_ordered <- matrix.lfc[order(match(rownames(matrix.lfc),
                                                     rownames(annotation_row))), , drop=FALSE]
#check again
rownames(matrix_row_ordered)
rownames(annotation_row) #yes!


#change order to matrix order 
matrix_row_ordered <- matrix.lfc[order(match(rownames(matrix.lfc),
                                                     rownames(annotation_row))), , drop=FALSE]

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.lfc))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row$Phylum %>% unique()
  length(Phyla) #14

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 14 colors for Phylum
display.brewer.pal(n = 14, name = 'Set3') #nice colors but we need more 


#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 14 #number of Phyla
color_phylum  <- colorRampPalette(brewer.pal(12, "Set3"))(colourCount) #vector with 14 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla, color_phylum)
  
 

#make colors for matrix
#Check first highest and ogest values in the matrix, in this case -4,+4
library(circlize)
function_color_matrix <- colorRamp2(c(-4, 0, 4), c("#018571", "white", "#A6611A")) #palette BrBG
function_color_matrix(seq(-4,4)) #gives you 8 colors with white in the middle 

# now plot the pheatmap
heatmap_ancombc_sig_all_lfc2_treatments <- ComplexHeatmap::pheatmap(matrix_row_ordered,
                         color= function_color_matrix,
                         name="LFC",
                         scale= "none", 
                         row_labels = species_ordered$Genus,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                        #annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         main = "Treatment",
                         fontsize_row = 15,
                         fontsize_col = 15)





```

##DA in WCS417 and CHA0, but not in WCS315

```{r}

#we will use the log2fold changes from ANCOM

ancombc_l

  
#Keep only the information needed (ASVs and Lfc)
lfc_417 <- ancombc_l$results$WCS417[,1, drop=FALSE]
lfc_315 <- ancombc_l$results$WCS315[,1, drop=FALSE] 
lfc_CHA0 <- ancombc_l$results$CHA0[,1, drop=FALSE] 


names(lfc_417) <- "lfc_WCS417"
names(lfc_315) <- "lfc_WCS315"
names(lfc_CHA0) <- "lfc_CHA0"

lfc_df <- cbind(lfc_417, lfc_315, lfc_CHA0) #merge

lfc_417_CHA0_df <- cbind(lfc_417, lfc_CHA0) #merge

dim(lfc_df) #6577
dim(lfc_417_CHA0_df) #6577

#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- rownames(ancombc_l$df_sig_asv_l$WCS417) #1000 ASVs
DA_WCS315_vector <- rownames(ancombc_l$df_sig_asv_l$WCS315) #1155 ASVs
DA_CHA0_vector <- rownames(ancombc_l$df_sig_asv_l$CHA0) #774 ASVs

#I want to keep only the ASVs  are differentially abundance under fo WCS417 and CHA0 to control, but not for WCS315

DA_417_CHA0_vector <- intersect(DA_WCS417_vector,DA_CHA0_vector) #447 ASVs
remove_ASV_vector <- intersect(DA_417_CHA0_vector, DA_WCS315_vector) #354 --> ASVs I wont to remove
DA_417_CHA0_vector <- outersect(DA_417_CHA0_vector, remove_ASV_vector) #93



DA_lfc_417_CHA0_df <- lfc_417_CHA0_df[DA_417_CHA0_vector, ] #matrix for pheatmap
length(DA_417_CHA0_vector) #93
dim(DA_lfc_417_CHA0_df) #93


#### Rhizosphere ####
matrix.lfc <- as.matrix(DA_lfc_417_CHA0_df)

#make ps with DA asvs
ps_DA <- prune_taxa(DA_417_CHA0_vector, ps_WW_rhizo_f)

#get species name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Genus")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Genus <- ifelse(is.na(df_row$Genus), df_row$asv_name, df_row$Genus)
df_row <- df_row %>%
           dplyr::select(Genus)


#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]))



#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.lfc)
rownames(annotation_row) #not in same order

#change order to matrix order 
matrix_row_ordered <- matrix.lfc[order(match(rownames(matrix.lfc),
                                                     rownames(annotation_row))), , drop=FALSE]
#check again
rownames(matrix_row_ordered)
rownames(annotation_row) #yes!

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.lfc))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row$Phylum %>% unique()
  length(Phyla) #12

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 30 colors for Phylum
display.brewer.pal(n = 12, name = 'Set3') #nice colors but we need more 

#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 12 #number of Phyla
color_phylum  <- brewer.pal(12, "Set3") #vector with 15 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla, color_phylum)


#define annotation colors
annotation_colors <- phylum_col
 

#make colors for matrix
#Check first highest and ogest values in the matrix, in this case -4,+4
library(circlize)
function_color_matrix <- colorRamp2(c(-4, 0, 4), c("#018571", "white", "#A6611A")) #palette BrBG
function_color_matrix(seq(-4,4)) #gives you 8 colors with white in the middle 

# now plot the pheatmap
heatmap_uniqueDA417CHA0 <- ComplexHeatmap::pheatmap(matrix_row_ordered,
                         color= function_color_matrix,
                         name="LFC",
                         scale= "none", 
                         row_labels = species_ordered$Genus,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         main = "Treatment",
                         fontsize_row = 15,
                         fontsize_col = 15)







##### if we use a cutoff of  lfc 2 ####

matrix.lfc <- as.matrix(DA_lfc_417_CHA0_df)

enriched <- matrix.lfc[apply(matrix.lfc > 2,1,any),] #10
depleated <- matrix.lfc[apply(matrix.lfc < -2,1,any),] #22

matrix.lfc2 <- rbind(enriched, depleated) #32

DA_417_CHA0__lfc2_v <- rownames(matrix.lfc2)

#make ps with DA asvs
ps_DA <- prune_taxa(DA_417_CHA0__lfc2_v, ps_WW_rhizo_f)

#get species name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Genus")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Genus <- ifelse(is.na(df_row$Genus), df_row$asv_name, df_row$Genus)
df_row <- df_row %>%
           dplyr::select(Genus)


#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]))



#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.lfc2)
rownames(annotation_row) #not in same order

#change order to matrix order 
matrix_row_ordered <- matrix.lfc2[order(match(rownames(matrix.lfc2),
                                                     rownames(annotation_row))), , drop=FALSE]
#check again
rownames(matrix_row_ordered)
rownames(annotation_row) #yes!

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.lfc2))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row$Phylum %>% unique()
  length(Phyla) #7

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 30 colors for Phylum
display.brewer.pal(n = 7, name = 'Set3') #nice colors but we need more 

#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 7 #number of Phyla
color_phylum  <- brewer.pal(7, "Set3") #vector with 15 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla, color_phylum)


#define annotation colors
annotation_colors <- phylum_col
 

#make colors for matrix
#Check first highest and ogest values in the matrix, in this case -4,+4
library(circlize)
function_color_matrix <- colorRamp2(c(-4, 0, 4), c("#018571", "white", "#A6611A")) #palette BrBG
function_color_matrix(seq(-4,4)) #gives you 8 colors with white in the middle 

# now plot the pheatmap
 heatmap_uniqueDA417CHA0_lfc2 <- ComplexHeatmap::pheatmap(matrix_row_ordered,
                         color= function_color_matrix,
                         name="LFC",
                         scale= "none", 
                         row_labels = species_ordered$Genus,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         #annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         main = "Treatment",
                         fontsize_row = 15,
                         fontsize_col = 15)



```


```{r}
#### Rhizosphere ####
matrix.log2 <- as.matrix(DA_log2fold_df)

#make ps with DA asvs
ps_DA <- prune_taxa(DA_taxa, ps_WW_rhizo_f)

#get Family name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Family")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Family <- ifelse(is.na(df_row$Family), df_row$asv_name, df_row$Family)
df_row <- df_row %>%
           dplyr::select(Family)


#change order DA_contrast_df
DA_contrast_df_ordered <- DA_contrast_df[order(match(rownames(DA_contrast_df),
                                                     rownames(annotation_row))), , drop=FALSE]
#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]),
    Contrast = DA_contrast_df_ordered)
 
#change it to factor so it can be colored
annotation_row$Contrast.DA_WCS417 <- as.factor(annotation_row$Contrast.DA_WCS417)
annotation_row$Contrast.DA_WCS315 <- as.factor(annotation_row$Contrast.DA_WCS315)
annotation_row$Contrast.DA_CHA0 <- as.factor(annotation_row$Contrast.DA_CHA0)


#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.log2)
rownames(annotation_row) #not in same order

#change order to matrix order 
matrix_row_ordered <- matrix.log2[order(match(rownames(matrix.log2),
                                                     rownames(annotation_row))), , drop=FALSE]
#check again
rownames(matrix_row_ordered)
rownames(annotation_row) #yes!

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.log2))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row %>% #17
         count(Phylum)

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 17 colors for Phylum
display.brewer.pal(n = 17, name = 'Set3') #nice colors but we need more 

#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 17 #number of Phyla
color_phylum  <- colorRampPalette(brewer.pal(12, "Set3"))(colourCount) #vector with 15 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla$Phylum, color_phylum)
  
 
#put all colors together
annotation_colors = list(
     #Phylum = phylum_col,
     Contrast.DA_WCS417 = c("1" = "black", "0" = "white"),
     Contrast.DA_WCS315 = c("1" = "black", "0" = "white"),
     Contrast.DA_CHA0 = c("1" = "black", "0" = "white"))

#make colors for matrix
library(circlize)
function_color_matrix <- colorRamp2(c(-7, 0, 7), c("#018571", "white", "#A6611A")) #palette BrBG
function_color_matrix(seq(-7,7)) #gives you 16 colors with white in the middle 

# now plot the pheatmap
ComplexHeatmap::pheatmap(matrix_row_ordered,
                         color= function_color_matrix,
                         name="L2FC",
                         scale= "none", 
                         row_labels = species_ordered$Family,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         #cluster_rows=TRUE,
                         main = "Treatment",
                         fontsize_row = 8,
                         fontsize_col = 8)



```

## Venn diagrams

```{r}
ancombc_l



#Keep only log2fold of DA taxa (from all the treatments pull)

DA_WCS417_vector <- rownames(ancombc_l$df_sig_asv_l$WCS417) #1000
DA_WCS315_vector <- rownames(ancombc_l$df_sig_asv_l$WCS315) #1155
DA_CHA0_vector <- rownames(ancombc_l$df_sig_asv_l$CHA0) #774

#With package  nVennR
library("nVennR")
library("rsvg")
library("grImport2")


myplot <- plotVenn(list(PsWCS417 = DA_WCS417_vector, 
           PsWCS315 = DA_WCS315_vector,
           PsCHA0 = DA_CHA0_vector),
          opacity = 0.4, borderWidth = 5,labelRegions = F, fontScale=3, 
          setColors = c("#A6CEE3", "#B2DF8A", "#CAB2D6"))


#Only when significant in all three and lfc>2



```



### What if we use only the robust ASVs

```{r}
#we will use the log2fold changes from normal deseq

deseq_norm_l

#Now, correct by shrienking with Log2 fold change 
  lfc_dds_417 <- lfcShrink(deseq_norm_l$result_test, coef=2, type="normal") 
  lfc_dds_315 <- lfcShrink(deseq_norm_l$result_test, coef=3, type="normal") 
  lfc_dds_CHA0 <- lfcShrink(deseq_norm_l$result_test, coef=4, type="normal") 
  
#Keep only the information needed (ASVs and log2fold)
log2fold_417 <- lfc_dds_417[2] 
log2fold_315 <- lfc_dds_315[2]
log2fold_CHA0 <- lfc_dds_CHA0[2]

names(log2fold_417) <- "log2fold_WCS417"
names(log2fold_315) <- "log2fold_WCS315"
names(log2fold_CHA0) <- "log2fold_CHA0"

log2fold_df <- cbind(log2fold_417, log2fold_315, log2fold_CHA0) #merge

dim(log2fold_df)

#Keep only log2fold of DA taxa (from all the treatments pull)


DA_WCS417_vector <- rownames(ps_robust_l$WCS417@tax_table)
DA_WCS315_vector <- rownames(ps_robust_l$wcs315@tax_table)
DA_CHA0_vector <- rownames(ps_robust_l$CHA0@tax_table)

DA_taxa <- c(DA_WCS417_vector, DA_WCS315_vector, DA_CHA0_vector) #join DA vectors
DA_taxa <-  unique(DA_taxa) #remove duplicate ASVs


DA_log2fold_df <- log2fold_df[DA_taxa, ] #matrix for pheatmap
dim(DA_log2fold_df)

#metadada

DA_contrast_df <- DA_log2fold_df 

DA_contrast_df <- DA_contrast_df %>% as.data.frame() %>% rownames_to_column(var = "ASV")

DA_contrast_df <- DA_contrast_df %>%
      mutate(DA_WCS417 = case_when(ASV %in% DA_WCS417_vector ~ "1",
                                TRUE ~ "0"))%>%
                mutate(DA_WCS315 = case_when(ASV %in% DA_WCS315_vector ~ "1",
                                TRUE ~ "0")) %>%
                mutate(DA_CHA0 = case_when(ASV %in% DA_CHA0_vector ~ "1",
                                TRUE ~ "0")) 


DA_contrast_df <- DA_contrast_df[, c(1, 5,6,7)]

DA_contrast_df <- DA_contrast_df %>%  column_to_rownames("ASV")




#### Rhizosphere ####
matrix.log2 <- as.matrix(DA_log2fold_df)

#make ps with DA asvs
ps_DA <- prune_taxa(DA_taxa, ps_WW_rhizo_f)

#get species name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Genus")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Genus <- ifelse(is.na(df_row$Genus), df_row$asv_name, df_row$Genus)
df_row <- df_row %>%
           dplyr::select(Genus)


#change order DA_contrast_df
DA_contrast_df_ordered <- DA_contrast_df[order(match(rownames(DA_contrast_df),
                                                     rownames(annotation_row))), , drop=FALSE]
#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]),
    Contrast = DA_contrast_df_ordered)
 
#change it to factor so it can be colored
annotation_row$Contrast.DA_WCS417 <- as.factor(annotation_row$Contrast.DA_WCS417)
annotation_row$Contrast.DA_WCS315 <- as.factor(annotation_row$Contrast.DA_WCS315)
annotation_row$Contrast.DA_CHA0 <- as.factor(annotation_row$Contrast.DA_CHA0)


#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.log2)
rownames(annotation_row) #not in same order

#change order to matrix order 
matrix_row_ordered <- matrix.log2[order(match(rownames(matrix.log2),
                                                     rownames(annotation_row))), , drop=FALSE]
#check again
rownames(matrix_row_ordered)
rownames(annotation_row) #yes!

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.log2))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row %>% #17
         count(Phylum)

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 17 colors for Phylum
display.brewer.pal(n = 17, name = 'Set3') #nice colors but we need more 

#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 17 #number of Phyla
color_phylum  <- colorRampPalette(brewer.pal(12, "Set3"))(colourCount) #vector with 15 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla$Phylum, color_phylum)
  
 
#put all colors together
annotation_colors = list(
     #Phylum = phylum_col,
     Contrast.DA_WCS417 = c("1" = "black", "0" = "white"),
     Contrast.DA_WCS315 = c("1" = "black", "0" = "white"),
     Contrast.DA_CHA0 = c("1" = "black", "0" = "white"))

#make colors for matrix
library(circlize)
function_color_matrix <- colorRamp2(c(-7, 0, 7), c("#018571", "white", "#A6611A")) #palette BrBG
function_color_matrix(seq(-7,7)) #gives you 16 colors with white in the middle 

# now plot the pheatmap
ComplexHeatmap::pheatmap(matrix_row_ordered,
                         color= function_color_matrix,
                         name="L2FC",
                         scale= "none", 
                         row_labels = species_ordered$Genus,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         #cluster_rows=TRUE,
                         main = "Treatment",
                         fontsize_row = 8,
                         fontsize_col = 20)

```

### Table with most importat ASVs

```{r}
matrix_row_ordered

#let's order the ASVs according to the ones with highest log2fold for WCS417
#change it to factor so it can be colored

DA_log2fold_df <- as.data.frame(DA_log2fold_df)
DA_log2fold_df <- DA_log2fold_df %>% arrange(desc(log2fold_WCS417))
matrix_row_ordered$log2fold_WCS417 <- as.numeric(matrix_row_ordered$log2fold_WCS417)
matrix_row_ordered <- matrix_row_ordered %>% arrange(desc(log2fold_WCS417))

str(matrix_row_ordered)





#Take our dataframe that tells as which ASV is signif DA under each treatment 
DA_contrast_df

#Add a new column that tell as under how many treatments the ASV is DA
DA_contrast_df$DA_WCS417 <- as.numeric(DA_contrast_df$DA_WCS417)
DA_contrast_df$DA_WCS315 <- as.numeric(DA_contrast_df$DA_WCS315)
DA_contrast_df$DA_CHA0 <- as.numeric(DA_contrast_df$DA_CHA0)

DA_contrast_df$n <-  rowSums(DA_contrast_df)

#Keep ASVs that are significant with 3 treatments
DA_contrast_all_sig_df <- filter(DA_contrast_df, n=="3")
dim(DA_contrast_all_sig_df) #17 ASVs are significant with 3 treatments

#Make a vector with the names of the 17 ASVs
all_sig_ASVs_v <- rownames(DA_contrast_all_sig_df)

#Check the log2fold of the 17 ASV
all_sig_DA_log2fold_df <- DA_log2fold_df[all_sig_ASVs_v,]

###Prepare for pheatmap###

matrix.log2 <- as.matrix(all_sig_DA_log2fold_df)

#make ps with DA asvs
ps_DA <- prune_taxa(all_sig_ASVs_v, ps_WW_rhizo_f)

#get species name if available, if not replace with name of ASV
df_row <- as.data.frame((tax_table(ps_DA))[, c("Genus")])
df_row <- df_row %>%
           mutate(asv_name=rownames(.)) #new column with rownames
df_row$Genus <- ifelse(is.na(df_row$Genus), df_row$asv_name, df_row$Genus)
df_row <- df_row %>%
           dplyr::select(Genus)


 #change order DA_contrast_df
DA_contrast_df_ordered <- DA_contrast_df[order(match(rownames(DA_contrast_df),
                                                     rownames(annotation_row))), , drop=FALSE]
#define annotation for rows
annotation_row = data.frame(
    Phylum = as.factor(tax_table(ps_DA)[, "Phylum"]))
 
#check if rownames are in same order 
rownames(tax_table(ps_DA)) #annotation row
rownames(df_row) #yes

#re-name rownames
rownames(annotation_row) = rownames(df_row)

#check if names are the same 
rownames(matrix.log2)
rownames(annotation_row) #not in same order

#make labels for rows 
df_row
species_ordered <- df_row[order(match(rownames(df_row), rownames(matrix.log2))), , drop = FALSE]

# select colors for factors

#### colors for Phyla ####

#count how many we have 
Phyla <- annotation_row %>% #5
         count(Phylum)

#select colors
library(RColorBrewer)
display.brewer.all() #gives you the palettes from RColorBrewer

#I like Set3 palette, let's get one with 17 colors for Phylum
display.brewer.pal(n = 5, name = 'Set3') #nice colors but we need more 

#Expand color palette with colorRampPalette
#https://www.datanovia.com/en/blog/easy-way-to-expand-color-palettes-in-r/
colourCount <- 5 #number of Phyla
color_phylum  <- colorRampPalette(brewer.pal(12, "Set3"))(colourCount) #vector with 15 colors
#make colors for phylum
phylum_col = as.data.frame(Phyla$Phylum, color_phylum)
  
  # now plot the pheatmap
ComplexHeatmap::pheatmap(matrix.log2,
                         color= function_color_matrix,
                         name="L2FC",
                         scale= "none", 
                         row_labels = species_ordered$Genus,
                         #annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         #annotation_colors = annotation_colors,
                         cluster_cols = FALSE,
                         cluster_rows = TRUE,
                         #cluster_rows=TRUE,
                         #main = "Treatment",
                         fontsize_row = 8,
                         fontsize_col = 10)


```

