---
title: "PCA"
author: "Zulema Carracedo Lorenzo"
date: "2023-06-05"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load needed packages

```{r}
library(DESeq2)
library(ggplot2)
library(apeglm)
library(glmGamPoi)
library(dplyr)
library(org.At.tair.db) #for functional enrichment

#for permanova
library(vegan)
library(pairwiseAdonis)

#tables
library(flextable)
library(officer)

```

#Data loading

```{r}
#Counts table
dat <- read.table('3_Arabidopsis_RNAseq_exp/data/salmon.merged.gene_counts.tsv', header=TRUE)
# Use gene_id as row index
rownames(dat) <- dat$gene_id
# Remove explicit gene_id and gene_name columns
countData <- subset(dat, select=-c(gene_id, gene_name)) 
# Round to integer counts: salmon (as used in e.g. nf-core/rnaseq) provides floats.
countData <- round(countData)
head(countData)

#metadata
#Sample info should describe experimental conditions for each sample in a 
#tabular format with samples as rows and variables as columns.
colData<- read.table('3_Arabidopsis_RNAseq_exp/data/metadata_organised.tsv', header=TRUE)
# By default we treat everything as categorical, i.e. a factor
#colData[, index] <- lapply(colData[, 1:ncol(colData)], as.factor) doesn't work
colData$Treatment <- as.factor(colData$Treatment)
colData$Time <- as.factor(colData$Time)
colData$Pseudomonas <- as.factor(colData$Pseudomonas)
# Use sample_id as row index
rownames(colData) <- colData$Sample
head(colData)
str(colData)

#Subset data by timepoint
colData_tp1 <- filter(colData, Time == "1")  #only metadata TP1
countData_tp1 <- select(countData, colData_tp1$Sample) #only count table TP1

colData_tp2 <- filter(colData, Time == "2")  #only metadata TP2
countData_tp2 <- dplyr::select(countData, colData_tp2$Sample) #only count table TP2

```



# 1. Differential expression testing
### LRT based testing

Create DESeq2 object, specify a full and reduced model, perform likelihood ratio test (LRT). For simple experiments (i.e. treatment VS control) this is the same as testing for a difference in means. For a more complex experiment (i.e. three genotypes and two timepoints), this is like doing an ANOVA.


```{r}
# Create DESeq2 object and specify full model
dds <- DESeqDataSetFromMatrix(
  countData = countData,
  colData = colData,
  design = ~ Pseudomonas + Time + Treatment
)

dds$Treatment <- relevel(dds$Treatment, "well_watered")

dds$Pseudomonas <- relevel(dds$Pseudomonas, "control")

# Perform LRT with reduced model
dds <- DESeq(
  dds,
  reduced = ~ Time,
  test = 'LRT',
  fitType = 'glmGamPoi'
)

# Scaled counts by DESeq2 size factor (will be used later for e.g. plotting)
scaledCountData <- as.data.frame(mapply('/', countData, dds$sizeFactor))
rownames(scaledCountData) <- rownames(countData)

# Select significant genes
alpha_cutoff <- 0.05
res <- results(dds, tidy = TRUE, alpha = alpha_cutoff)
selection <- res$padj < alpha_cutoff
selection[is.na(selection)] <- FALSE # Zero-expression genes will have NAs, do not select
sig <- res[selection, ]

# Sort by adjusted p-value
sig <- sig[order(sig$padj),]
sig
```


### PCA on DESeq2 normalized counts

This uses a variance stabilizing transformation: either `rlog` or `vst`.
Use `rlog` for small to medium sized experiments, and `vst` for big experiments (Or wait a long time). 

```{r}
#PCA on DESeq2 normalized counts

PCA <-plotPCA(
  rlog(dds, blind=FALSE, fitType='glmGamPoi'),
  intgroup=c('Pseudomonas','Treatment', 'Time'),
  returnData=TRUE)

percentVar <- round(100 * attr(PCA, "percentVar"))

PCA$Pseudomonas <- factor(PCA$Pseudomonas, c("control", "WCS417", "WCS315", "CHA0"))

  ggplot(PCA, aes(PC1, PC2, fill=Pseudomonas, shape=Treatment, color = Time)) +
  geom_point(size=8) +
  scale_shape_manual(values = c(21, 22) )+
  scale_color_manual(values=c("#1F78B4","#E31A1C"))+
  scale_fill_manual(values = c("#FFFF99", "#A6CEE3", "#B2DF8A", "#CAB2D6")) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  guides(fill=guide_legend(override.aes=list(shape=21)))+
  #guides(color=guide_legend(override.aes=list(shape=21))) +
  #guides(fill=guide_legend(override.aes=list(shape=21,fill=c("#FFFF99","#A6CEE3","#B2DF8A","#CAB2D6"),color=c("#1F78B4","#E31A1C")))) +
  theme_bw(18) +
  labs(fill = "Pseudomonas strain", color = "Time", shape = "Water regime") +
  #guides(fill= "none") +
  theme(axis.line = element_line(colour = "black"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          panel.background = element_blank()) 
  

  PCA # data frame from whch I want to make a PCA plot
  #rename water regime names and pseudomonas strains for figures
  

  PCA <- PCA %>% mutate('Pseudomonas_strain' =
                          case_when(Pseudomonas== "control" ~ "Control",
                                    Pseudomonas=="WCS417" ~ "PsWCS417",
                                    Pseudomonas=="WCS315" ~ "PsWCS315",
                                    Pseudomonas=="CHA0" ~ "PsCHA0"))
                                    
  colnames(PCA) <- c("PC1", "PC2", "group", "Pseudomonas", "water_regime", "Time_point", "name", "Pseudomonas_strain") # Time -> timepoint
  
  PCA$Pseudomonas_strain <- factor(PCA$Pseudomonas_strain, c("Control", "PsWCS417", "PsWCS315", "PsCHA0"))
  PCA$water_regime <- factor(PCA$water_regime, c("well_watered", "drought"))
  
#make PCA
  ggplot(PCA, aes(PC1, PC2, fill=Pseudomonas_strain, color= water_regime, shape = Time_point))+
    geom_point(size=8, stroke = 2)+
    scale_shape_manual(values = c(21, 22) )+
    scale_color_manual(values=c("black","grey"))+
    scale_fill_manual(values = c("#76CFE8", "#C9DC53", "#F99F58", "#9771A5")) +
    guides(fill=guide_legend(override.aes=list(shape=21)))+
    #geom_mark_ellipse(expand = 0.03,aes(fill=Pseudomonas_strain)) +
    xlab(paste0("PC1: ",percentVar[1],"% variance")) +
    ylab(paste0("PC2: ",percentVar[2],"% variance")) +
    theme_bw(18) +
    labs(fill = "Pseudomonas strain", color = "Water regime", shape = "Time point") +
    theme(axis.line = element_line(colour = "black"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          panel.background = element_blank()) 
  
  #save it
  mapply(function(x)
    ggsave(
      plot = last_plot(),
      filename =x,
      path = "./3_Arabidopsis_RNAseq_exp/results/plots",
      scale = 1.8,
      width = 200,
      height = 100,
      units = "mm",
      dpi = 300),
    x=c("Fig.2A_PCA_RNAseq.svg","Fig.2A_PCA_RNAseq.png"))
  

```

#PERMANOVA
```{r}
pca_dist <- dist(PCA[, c("PC1", "PC2")])  # Euclidean distance on PC scores

# Run PERMANOVA (replace factors with yours)
adonis_result <- adonis2(
  pca_dist ~ Pseudomonas_strain + Time_point  + water_regime ,
  data = PCA,
  permutations = 999,
  method = "euclidean"
)

print(adonis_result)

#save table in word
adonis_df <- as.data.frame(adonis_result)

# Add row names as a proper column
adonis_df$Factor <- rownames(adonis_df)

# Move 'Factor' to the first column
adonis_df <- adonis_df[, c("Factor", setdiff(names(adonis_df), "Factor"))]

# Round values for clarity
adonis_df$`Pr(>F)` <- formatC(adonis_df$`Pr(>F)`, format = "f", digits = 3)
adonis_df$R2 <- round(adonis_df$R2, 3)
adonis_df$F <- round(adonis_df$F, 2)
adonis_df$SumOfSqs <- round(adonis_df$SumOfSqs, 2)

# Create styled flextable
adonis_ft <- flextable(adonis_df) |>
  theme_booktabs() |>
  autofit() |>
  bold(i = ~ `Pr(>F)` < 0.05, part = "body") |>
  set_header_labels(
    Factor = "Factor", Df = "DF", SumOfSqs = "Sum Sq",
    R2 = "R²", F = "F-value", `Pr(>F)` = "p-value"
  )

#export in a word
doc <- read_docx() |>
  body_add_par("Table: PERMANOVA Results on PCA Scores", style = "heading 2") |>
  body_add_flextable(adonis_ft)

print(doc, target = "PERMANOVA_results.docx")



```

##Pairwise compaisons
```{r}

# Assuming pca_dist is your distance matrix and pcaData contains the grouping factor
pairwise_results <- pairwise.adonis2(
  pca_dist ~ Pseudomonas_strain + Time_point  + water_regime,
  data = PCA,
  permutations = 999,
  p.adjust.m = "bonferroni"  # or "fdr" for false discovery rate
)

print(pairwise_results)


#save results for word table


# Create a new Word document
doc <- read_docx()

# Loop through each comparison in the pairwise results
for (comparison in names(pairwise_results)) {
  
  if (comparison != "parent_call") {
    df <- as.data.frame(pairwise_results[[comparison]])
    
    # Add Factor as a column (if rownames are not already in a column)
    df$Factor <- rownames(df)
    df <- df[, c("Factor", setdiff(names(df), "Factor"))]
    
    # Round values for clarity
    num_cols <- sapply(df, is.numeric)
    df[num_cols] <- lapply(df[num_cols], function(x) round(x, 3))
    
    # Create flextable
    ft <- flextable(df) %>%
      autofit() %>%
      theme_booktabs() %>%
      bold(i = ~ `Pr(>F)` < 0.05, part = "body") %>%
      set_header_labels(
        Factor = "Factor", Df = "DF", SumOfSqs = "Sum Sq",
        R2 = "R²", F = "F-value", `Pr(>F)` = "p-value"
      )
    
    # Add section title and table to document
    doc <- doc %>%
      body_add_par(paste("Comparison:", comparison), style = "heading 2") %>%
      body_add_flextable(ft) %>%
      body_add_par("", style = "Normal")  # spacing
  }
}

# Save document
print(doc, target = "pairwise_PERMANOVA_results.docx")

```

```
